---
title: "Personal diet recommendation"
subtitle: "Supplementary Materials"
author:
- JARI TURKIA \textsuperscript{1,2}
- URSULA SCHWAB \textsuperscript{3,4}
- VILLE HAUTAMÄKI \textsuperscript{1,5}
- \textsuperscript{1} School of Computing, University of Eastern Finland, 80101 Joensuu, Finland
- \textsuperscript{2} CGI Suomi, Joensuu, Finland
- \textsuperscript{3} School of Medicine, Institute of Public Health and Clinical Nutrition,
- University of Eastern Finland, Kuopio, Finland
- \textsuperscript{4} Department of Medicine, Endocrinology and Clinical Nutrition,
- Kuopio University Hospital, Kuopio, Finland
- \textsuperscript{5} Department of Electrical and Computer Engineering, 
- National University of Singapore, Singapore

output:
  pdf_document:
    latex_engine: xelatex
    fig_caption: yes
    includes:
      in_header: custom_figure_captions.tex
  html_document: default
header-includes: \usepackage{float}  
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra) # for HTML and Latex tables

knitr::opts_chunk$set(echo = TRUE, fig.align="center", fig.pos = 'H')

# this allows using tikz rendering for plots with "dev=tikz"
knit_hooks$set(plot = function(x, options) {
  if ('tikz' %in% options$dev && !options$external) {
    hook_plot_tex(x, options)
  } else hook_plot_md(x, options)
})

# Fix seed for random number generator for getting consistent results in kmeans etc.
fixed_seed <- 678

# Load common MEBN package
source("mebn/v2/MEBNv2.r")

# Online version (TRUE) uses color images while (FALSE) renders exact figures for the article 
RenderArticleFigures <- FALSE
```

This notebook reproduces in detail the analysis of personal diet recommendations described in the main
article.

The notebook execution generates all the figures and tables that are included in the article and produces also the referenced supplementary figures. The article is accompanied by a PDF rendition of the notebook that shows all the supplementary figures and important parts of the program code so that the analysis can be followed in detail. The executable RMarkdown notebook with data can be found in a public GitHub repository of the corresponding author.

# Dialysis patient data

```{r data_loading, echo=FALSE, message=FALSE, warning=FALSE}
library(lubridate)
library(dplyr)
library(tidyr)

# Read the data description
datadesc_fat_epros <- read.csv(file="data/DIALYSIS_data_description_fat_epros.csv", header = TRUE, sep = ";")

# Read the actual data matching the description
dialysis <- read.csv(file="data/DIALYSIS.csv", sep=";", dec=",")

# Define factors
dialysis$potilas <- factor(dialysis$potilas)
dialysis$nesterajoitus <- factor(dialysis$nesterajoitus)
dialysis$verenpainelaakitys <- factor(dialysis$verenpainelaakitys)
dialysis$verenrasvojen_laakitys <- factor(dialysis$verenrasvojenlaakitys)
dialysis$mielialalaakitys <- factor(dialysis$mielialalaakitys)
dialysis$fosforinsitoja <- factor(dialysis$fosforinsitoja)
dialysis$kaliuminsitoja <- factor(dialysis$kaliuminsitoja)
dialysis$anemian_hoito <- factor(dialysis$anemianhoito)
dialysis$diabeteslaakitys <- factor(dialysis$diabeteslaakitys)
dialysis$antibiootti <- factor(dialysis$antibiootti)
dialysis$marevan <- factor(dialysis$marevan)
dialysis$nesteenpoisto <- factor(dialysis$nesteenpoisto)
dialysis$ummetuslaake <- factor(dialysis$ummetuslaake)
dialysis$verenohennus <- factor(dialysis$verenohennus)
dialysis$akt_dvit <- factor(dialysis$aktdvit)
dialysis$renavit <- factor(dialysis$renavit)
dialysis$cad <- factor(dialysis$cad)

# Calculate missing E% for fats 
# https://fineli.fi/fineli/fi/ravintotekijat/2331

dialysis$mufaepros <- dialysis$mufa * 9 / dialysis$energiakcal * 100 # Rasvan energiakerroin on 37 kJ/g (9 kcal/g)
dialysis$pufaepros <- dialysis$pufa * 9 / dialysis$energiakcal * 100 # Rasvan energiakerroin on 37 kJ/g (9 kcal/g)
#dialysis$safaepros2 <- dialysis$safa * 9 / dialysis$energiakcal * 100 # Rasvan energiakerroin on 37 kJ/g (9 kcal/g)

# Add these new E% features as predictors
mufa_e <- datadesc_fat_epros[datadesc_fat_epros$Name=="mufa",]
mufa_e$Name <- "mufaepros"
mufa_e$Unit <- "E%"
mufa_e$Description <- "Monounsaturated Fatty Acids, E%"
mufa_e$Descriptionfin <- "Mufa E%"
datadesc_fat_epros <- rbind(datadesc_fat_epros, mufa_e)

pufa_e <- datadesc_fat_epros[datadesc_fat_epros$Name=="pufa",]
pufa_e$Name <- "pufaepros"
pufa_e$Unit <- "E%"
pufa_e$Description <- "Polyunsaturated Fatty Acids, E%"
pufa_e$Descriptionfin <- "Pufa E%"
datadesc_fat_epros <- rbind(datadesc_fat_epros, pufa_e)

# Remove g/d features from predictors
datadesc_fat_epros[datadesc_fat_epros$Name=="mufa",]$Order <- 0
datadesc_fat_epros[datadesc_fat_epros$Name=="pufa",]$Order <- 0

# gender: female = 1, male = 0
dialysis$sukupuoli <- factor(ifelse(dialysis$sukupuoli == "nainen", 1, 0))

# Days between laboratory test and food record interview
dialysis$responsedays <- difftime(as.Date(dialysis$labraaika, '%d.%m.%Y'), as.Date(dialysis$ravhaastaika, '%d.%m.%Y'), units = c("days"))

responsedays_desc <- datadesc_fat_epros[datadesc_fat_epros$Name=="ika",]
responsedays_desc$Order <- 0 # OMITTED !!
responsedays_desc$Distribution <- "Integer"  # Filter this out from normalization etc.
responsedays_desc$Name <- "responsedays"
responsedays_desc$Unit <- "days"
responsedays_desc$Description <- "responsetime"
responsedays_desc$Descriptionfin <- "vasteaika"
datadesc_fat_epros <- rbind(datadesc_fat_epros, responsedays_desc)

# important: dataset is ordered by successive patients and observations so that estimation works correctly
dialysis <- dialysis[order(dialysis$potilas, dialysis$havainto),]

# remove patients with one observation only

# dialysis
patients_with_two_obs <- dialysis %>% 
  group_by(potilas) %>% 
  summarise(havainto_sum = sum(havainto)) %>%
  filter(havainto_sum == 3) %>%
  select(potilas) %>%
  unlist() %>% as.vector()

dialysis <- dialysis[dialysis$potilas %in% patients_with_two_obs,]

# - reset patient levels
dialysis$potilas <- factor(dialysis$potilas)

# Define how to iterate through the graph
# - same targets
assumedtargets <- datadesc_fat_epros[datadesc_fat_epros$Order==200,]
```

```{r two_sets_of_predictors, echo=FALSE, message=FALSE}
# - different sets of predictors

## FAT EPROS
assumedpredictors_fat_epros <- datadesc_fat_epros[datadesc_fat_epros$Order==100,]

# Ordering for table
assumedpredictors_fat_epros <- assumedpredictors_fat_epros[order(assumedpredictors_fat_epros$Unit, assumedpredictors_fat_epros$Description),]

gaussian_preds <- as.vector(assumedpredictors_fat_epros[assumedpredictors_fat_epros$Distribution == "Gaussian" | assumedpredictors_fat_epros$Distribution == "LogNormal",]$Name)
show_decimals <- as.vector(assumedpredictors_fat_epros[assumedpredictors_fat_epros$Distribution == "Gaussian" | assumedpredictors_fat_epros$Distribution == "LogNormal",]$Decimals)

pred_table <- as.data.frame(assumedpredictors_fat_epros[assumedpredictors_fat_epros$Distribution == "Gaussian" | assumedpredictors_fat_epros$Distribution == "LogNormal",]$Description, row.names = NULL)

range_mat <- paste0(round(apply(t(dialysis[gaussian_preds]), 1, function(x) mean(x)),show_decimals), " (",round(apply(t(dialysis[gaussian_preds]), 1, function(x) min(x)),show_decimals)," - ",round(apply(t(dialysis[gaussian_preds]), 1, function(x) max(x)),show_decimals),") ", as.vector(assumedpredictors_fat_epros[assumedpredictors_fat_epros$Distribution == "Gaussian",]$Unit))

prior_range <- paste0(assumedpredictors_fat_epros[assumedpredictors_fat_epros$Distribution == "Gaussian",]$Lowerbound, " - ", assumedpredictors_fat_epros[assumedpredictors_fat_epros$Distribution == "Gaussian",]$Upperbound, " ", assumedpredictors_fat_epros[assumedpredictors_fat_epros$Distribution == "Gaussian",]$Unit)

pred_table_fat_epros <- cbind(pred_table, range_mat)
colnames(pred_table_fat_epros) <- c("Nutrient", "Sample avg. (min-max)")

```

```{r table1, echo=FALSE, message=FALSE, warning=FALSE}
library(kableExtra)

# The analyzed dataset consists of food records and laboratory measurements from end-stage renal patients in dialysis. The following nutrients are considered possible predictors of 
# concentrations. Note that we use energy-% of fats and protein as a unit in our analysis. This is referred to as Table 1 in the main article.

# Latex notation from tables/nutrient_table.tex is used in the article

phos_id <- which(pred_table_fat_epros$Nutrient == "Phosphorous")
pot_id <- which(pred_table_fat_epros$Nutrient == "Potassium")

table1 <- kable(pred_table_fat_epros, booktabs = T, row.names = NA, format = "latex", caption = "Nutrient predictors of the model.") %>%
  kable_styling(full_width = TRUE) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  row_spec(0,bold=TRUE) %>% 
  row_spec(phos_id,bold=TRUE) %>%
  row_spec(pot_id,bold=TRUE)

#table1

table1 %>% save_kable(file = "tables/nutrient_table.pdf", keep_tex = TRUE)
```
```{r table2, echo=FALSE, message=FALSE}
library(kableExtra)

# The following personal details and medication are considered as predictors. This is referenced as Table 2 in the article.

medication_preds <- as.vector(assumedpredictors_fat_epros[assumedpredictors_fat_epros$Distribution == "Bernoulli",]$Name)
medication_table <- as.data.frame(assumedpredictors_fat_epros[assumedpredictors_fat_epros$Distribution == "Bernoulli",]$Description, row.names = NULL)

# Add dialysis treatments to table 
medication_preds <- c(medication_preds, "KHD", "OHD", "PD")
medication_table <- rbind(medication_table, "Home hemodialysis", "Hospital hemodialysis", "Peritoneal dialysis")

# Hot-encode the treatments in data (PD/KHD/OHD)
dialysis$PD <- 0  
dialysis$KHD <- 0  
dialysis$OHD <- 0  
dialysis[dialysis$hoitomuoto == "PD",]$PD <- 1
dialysis[dialysis$hoitomuoto == "KHD",]$KHD <- 1
dialysis[dialysis$hoitomuoto == "OHD",]$OHD <- 1

medication_table <- cbind(medication_table, paste0(round(colSums(dialysis[medication_preds] == 1) / nrow(dialysis[medication_preds]) * 100, 0), "%"))
colnames(medication_table) <- c("Personal detail", "Percentage of patients")

# fix gender description in table
gender_row <- medication_table[medication_table[1]=="Gender",]
gender_row[2] <- paste0(gender_row[2], " female")

medication_table[medication_table[1]=="Gender",] <- gender_row

# Latex notation from tables/personinfo_table.tex is used in the article

table2 <- kable(medication_table, booktabs = T, row.names = NA, format = "latex", caption = "Personal details that are used as predictors.") %>%
  kable_styling(full_width = TRUE) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  row_spec(0,bold=TRUE)

#table2

table2 %>% save_kable(file = "tables/personinfo_table.pdf", keep_tex = TRUE)
```
Let us explore then how the intake of dietary potassium and phosphorus correspond in data to plasma concentrations of potassium, phosphate and albumin.

```{r scatterplot, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.align="left", fig.width=6, fig.cap="Figure shows the progress between two observations of plasma potassium (P-K), fasting plasma phosphate (fP-Pi) and plasma albumin (P-Alb) concentrations of the studied patients. White regions in the panels show the recommended concentration levels, P-K 3.4 - 4.7 mmol/l, fP-Pi < 1.8 mmol/l and P-Alb 34 - 48 g/l. Vertical black lines denote the commonly recommended maximum intakes of these nutrients. The goal is to find personal intake levels that keep the concentrations within recommended normal ranges, if possible. The figure is plotted with ggplot2 package for R language (v 3.3.5, https://ggplot2.tidyverse.org)."}
library(ggplot2)
library(gridExtra)
library(grid)

pk_k_plot <- ggplot() +
  geom_point(data=dialysis, aes(y = pk, x=kalium, color=potilas), size=0.4, show.legend = FALSE) +
  geom_line(data=dialysis, aes(y = pk, x=kalium, color=potilas), size=0.4, show.legend = FALSE) +
  #geom_text(data=dialysis, aes(y = pk, x=kalium, label=potilas)) +
  geom_vline(xintercept = 2500, linetype="solid", color = "black", size=0.5) +
  labs(x = "Potassium (mg/d)", y = "P-K (mmol/l)") + 
  scale_x_continuous() +
  scale_y_continuous(breaks=c(3,3.4,4,4.7,5,6)) +
  annotate("rect",ymin=-Inf, ymax=c(3.4), xmin=-Inf, xmax=Inf, alpha = .2) + 
  annotate("rect",ymin=c(4.7), ymax=Inf, xmin=-Inf, xmax=Inf, alpha = .2) + 
  theme_bw() +
  theme(axis.title.x = element_blank(), axis.text = element_text(size=6), axis.title = element_text(size=8)) 

pk_p_plot <- ggplot() +
  geom_point(data=dialysis, aes(y = pk, x=fosfori, color=potilas), size=0.4, show.legend = FALSE) +
  geom_line(data=dialysis, aes(y = pk, x=fosfori, color=potilas), size=0.4, show.legend = FALSE) +
  #geom_text(data=dialysis, aes(y = pk, x=fosfori, label=potilas)) +
  geom_vline(xintercept = 1000, linetype="solid", color = "black", size=0.5) +
  labs(x = "Phosphorus (mg/d)", y = "P-K (mmol/l)") + 
  scale_x_continuous() +
  scale_y_continuous(breaks=c(3,3.4,4,4.7,5,6)) +
  annotate("rect",ymin=-Inf, ymax=c(3.4), xmin=-Inf, xmax=Inf, alpha = .2) + 
  annotate("rect",ymin=c(4.7), ymax=Inf, xmin=-Inf, xmax=Inf, alpha = .2) + 
  theme_bw() +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text = element_text(size=6), axis.title = element_text(size=8)) 

fppi_k_plot <- ggplot() +
  geom_point(data=dialysis, aes(y = fppi, x=kalium, color=potilas), size=0.4, show.legend = FALSE) +
  geom_line(data=dialysis, aes(y = fppi, x=kalium, color=potilas), size=0.4, show.legend = FALSE) +
  geom_vline(xintercept = 2500, linetype="solid", color = "black", size=0.5) +
  scale_x_continuous() +
  scale_y_continuous(breaks=c(1,1.8,2,3)) +
  annotate("rect",ymax=Inf, ymin=c(1.8), xmin=-Inf, xmax=Inf, alpha = .2) + 
  labs(x = "Potassium (mg/d)", y = "fP-Pi (mmol/l)") + 
  theme_bw() +
  theme(axis.title.x = element_blank(), axis.text = element_text(size=6), axis.title = element_text(size=8))

fppi_p_plot <- ggplot() +
  geom_point(data=dialysis, aes(y = fppi, x=fosfori, color=potilas), size=0.4, show.legend = FALSE) +
  geom_line(data=dialysis, aes(y = fppi, x=fosfori, color=potilas), size=0.4, show.legend = FALSE) +
  geom_vline(xintercept = 1000, linetype="solid", color = "black", size=0.5) +
  scale_x_continuous() +
  scale_y_continuous(breaks=c(1,1.8,2,3)) +
  annotate("rect",ymax=Inf, ymin=c(1.8), xmin=-Inf, xmax=Inf, alpha = .2) + 
  labs(x = "Phosphorus (mg/d)", y = "fP-Pi (mmol/l)") + theme(axis.title.x = element_blank()) +
  theme_bw() +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text = element_text(size=6), axis.title = element_text(size=8)) 

palb_k <- ggplot() +
  geom_point(data=dialysis, aes(y = palb, x=kalium, color=potilas), size=0.4, show.legend = FALSE) +
  geom_line(data=dialysis, aes(y = palb, x=kalium, color=potilas), size=0.4, show.legend = FALSE) +
  #geom_text(data=dialysis, aes(y = pk, x=kalium, label=potilas)) +
  geom_vline(xintercept = 2500, linetype="solid", color = "black", size=0.5) +
  labs(x = "Potassium (mg/d)", y = "P-Alb (g/l)") +
  scale_x_continuous() +
  scale_y_continuous(breaks=c(34,48)) +
  annotate("rect",ymin=-Inf, ymax=c(34), xmin=-Inf, xmax=Inf, alpha = .2) + 
  annotate("rect",ymin=c(48), ymax=Inf, xmin=-Inf, xmax=Inf, alpha = .2) + 
  theme_bw() +
  theme(axis.text = element_text(size=6), axis.title = element_text(size=8)) 

palb_p <- ggplot() +
  geom_point(data=dialysis, aes(y = palb, x=fosfori, color=potilas), size=0.4, show.legend = FALSE) +
  geom_line(data=dialysis, aes(y = palb, x=fosfori, color=potilas), size=0.4, show.legend = FALSE) +
  geom_vline(xintercept = 1000, linetype="solid", color = "black", size=0.5) +
  scale_x_continuous() +
  scale_y_continuous(breaks=c(34,48)) +
  annotate("rect",ymin=-Inf, ymax=c(34), xmin=-Inf, xmax=Inf, alpha = .2) + 
  annotate("rect",ymin=c(48), ymax=Inf, xmin=-Inf, xmax=Inf, alpha = .2) + 
  labs(x = "Phosphorus (mg/d)", y = "P-Alb (g/l)") +
  theme_bw() +
  theme(axis.title.y = element_blank(), axis.text = element_text(size=6), axis.title = element_text(size=8)) 

p <- grid.arrange(pk_k_plot, pk_p_plot, fppi_k_plot, fppi_p_plot, palb_k, palb_p, nrow = 3, ncol=2, padding=0)
#ggsave("figures/data_scatterplot.pdf", plot = p, width = 6, height = 7, scale = 1)
```

Aim of this analysis is to find such personal levels of potassium and phosphorous intakes that keep all these concentrations in their recommended levels marked with white regions, if possible. 

# Development of nutrition reaction model

In this work we construct personal generative models for concentrations that allow conditioning personal diet recommendations. Essential part of these models is the personal reactions to nutrients and other predictors. Simultaneous reactions on all considered concentrations are modeled as multivariate model that has all the concentrations are response variables.

For comparison, we estimate multivariate systems with and without cross-model covariance. With cross-model covariance estimated, the model corresponds to seemingly unrelated model system, and without the model is Bayesian network with separated local distributions. This more simple model is estimated first with only potassium and phosphorous concentrations as responses.

```{r model_separate_univariate_preparation, echo=FALSE, eval=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
library(rstan)
source("mebn/v2/MEBNv2.r")

rstan_options (auto_write=TRUE)
options (mc.cores=parallel::detectCores ()) 
```

```{r model_separate_univariate, echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
initial_graph <- mebn.fully_connected_bipartite_graph(datadesc_fat_epros)

pk_fppi_targets <- datadesc_fat_epros[datadesc_fat_epros$Name %in% c('pk','fppi'),]

dialdiet_gamma <- mebn.bipartite_model(reaction_graph = initial_graph, 
                                   inputdata = dialysis,
                                   predictor_columns = assumedpredictors_fat_epros, 
                                   assumed_targets = pk_fppi_targets, 
                                   group_column = "potilas",
                                   local_estimation = mebn.sampling,
                                   local_model_cache = "models/BLMM_gamma_separate", 
                                   stan_model_file = "mebn/v2/BLMM_gamma.stan",
                                   normalize_values = TRUE)

# Write the generated graph in a GraphML file
write.graph(dialdiet_gamma, "graphs/dialysis_gamma_separate.graphml", "graphml")
```

In Bayesian network both responses were estimated separately. Next, we estimate a multivariate model where both distribution are estimated during single sampling. It does not factorize into separate distributions, but allows using more data.

```{r graph_with_gamma_qr_mv_epros_prep, echo=FALSE, warning=FALSE, message=FALSE}
initial_graph <- mebn.fully_connected_bipartite_graph(datadesc_fat_epros)

pk_fppi_targets <- datadesc_fat_epros[datadesc_fat_epros$Name %in% c('pk','fppi'),]
```


```{r graph_with_gamma_qr_mv_epros, echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
dialdiet_gamma_mv2_epros <- mebn.bipartite_multivariate(reaction_graph = initial_graph, 
                                   inputdata = dialysis,
                                   predictor_columns = assumedpredictors_fat_epros, 
                                   assumed_targets = pk_fppi_targets, 
                                   group_column = "potilas",
                                   local_estimation = mebn.multivariate_sampling,
                                   local_model_cache = 
                                     "models/BLMM_gamma_qr_multivariate2/fat_epros", 
                                   stan_model_file = "mebn/v2/BLMM_gamma_qr_mv.stan",
                                   normalize_values = TRUE)

write.graph(dialdiet_gamma_mv2_epros, 
            "graphs/dialysis_gamma_multivariate2_epros.graphml", "graphml")
```
Next we like to add plasma albumin concentration as a third constraint in the model, but unfortunately 8 of 37 patients have missing albumin measurements. From such a small dataset we don't want to remove any patients and so we predict the missing albumin levels and impute them to dataset. 

For prediction, we estimate the model without missing concentrations in P-Alb. The rows with missing values are held out from the density estimation, but patients are kept in the model for estimating their parameters.

```{r graph_with_gamma_qr_mv3_epros, echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
pk_fppi_palb_targets <- 
  datadesc_fat_epros[datadesc_fat_epros$Name %in% c('pk','fppi','palb'),]

# 0/1-index for palb = NA
holdout_index <- as.vector(as.numeric(is.na(dialysis$palb)))

# Stan does not support NA in data (in Y), so let's change NA to magic number
if (anyNA(dialysis$palb)) dialysis[is.na(dialysis$palb),]$palb <- -1

initial_graph <- mebn.fully_connected_bipartite_graph(datadesc_fat_epros)

dialdiet_gamma_mv3_missing_palb <- mebn.bipartite_multivariate(
                                   reaction_graph = initial_graph, 
                                   inputdata = dialysis,
                                   targetdata = holdout_index,
                                   predictor_columns = assumedpredictors_fat_epros, 
                                   assumed_targets = pk_fppi_palb_targets, 
                                   group_column = "potilas",
                                   local_estimation = mebn.multivariate_sampling,
                                   local_model_cache = 
                                     "models/BLMM_gamma_qr_multivariate3/fat_epros_missing_palb", 
                                   stan_model_file = "mebn/v2/BLMM_gamma_qr_mv_cv.stan",
                                   normalize_values = TRUE)

```
# Personal data imputation models

Now we can create personal models for patients that have missing plasma albumin measurements. These models are then used for making personal predictions for replacing missing values.

```{r extract_personal_generative_models1, echo=TRUE, eval=TRUE, warning=FALSE, cache=TRUE, eval=FALSE}
# Extract personal generative models for patients who have missing P-Alb values

patients_with_missing_palb <- unique(as.vector(dialysis[is.na(dialysis$palb),]$potilas))

# - initial graph structure
initial_graph <- mebn.fully_connected_bipartite_graph(datadesc_fat_epros)

# - previously estimated graphical model with all the persons
local_distributions <- pk_fppi_palb_targets
local_distributions$modelcache <- 
  "models/BLMM_gamma_qr_multivariate3/fat_epros_missing_palb"

# - get personal data, normalized and original

# - statistics for vertex levels are calculated from normalized data
predictors <- nrow(assumedpredictors_fat_epros)
normalized_input <- sapply(1:predictors, mebn.scale_gaussians, 
                           data = dialysis, datadesc = assumedpredictors_fat_epros)
normalized_input_df <- as.data.frame(normalized_input)

datadesc <- datadesc_fat_epros

```

```{r extract_personal_generative_models2, echo=FALSE, eval=TRUE, warning=FALSE, cache=TRUE, eval=FALSE}
for (person_id in 1:length(patients_with_missing_palb)) {
  
  # - pick rows for the selected person
  subject_code <- patients_with_missing_palb[person_id]
  personal_data_df <- cbind(dialysis$potilas, normalized_input_df)
  personal_data_df <- personal_data_df[personal_data_df$`dialysis$potilas` == subject_code,]
  personal_data <- as.matrix(subset(personal_data_df, select = -c(`dialysis$potilas`)))
  
  # - store also these original stats in graph
  personal_data_org <- subset(dialysis[dialysis$potilas == subject_code,], 
                              select = as.vector(assumedpredictors_fat_epros$Name))
  
  personal_model_dir <- paste0("data_imputation_models/", subject_code)
  
  # Generate a personal graph in directory
  # - This function extracts the personal model from a single level multivariate model
  
  personal_graph <- mebn.extract_personal_graph_from_mv(person_id, initial_graph, 
                    personal_model_dir, assumedpredictors_fat_epros, pk_fppi_palb_targets, 
                    "models/BLMM_gamma_qr_multivariate3/fat_epros_missing_palb", 
                    personal_data, personal_data_org, datadesc_fat_epros)
}
```

Then we predict missing values with these data imputing models

```{r, palb_imputation1, echo=FALSE, eval=FALSE}
library(igraph)
source("mebn/v3/MEBNv3.r")

# this vector is imputed with predicted values
original_palb <- dialysis$palb
imput_length <- length(original_palb)

# Normalize input for predictions (as it was in the likelihood estimation)
predictors <- nrow(assumedpredictors_fat_epros)
normalized_input <- sapply(1:predictors, mebn.scale_gaussians, 
                           data = dialysis, datadesc = assumedpredictors_fat_epros)
normalized_input_df <- as.data.frame(normalized_input)

normalized_input_df <- cbind(dialysis$potilas, normalized_input_df)
normalized_input_df <- cbind(dialysis$palb, normalized_input_df)

colnames(normalized_input_df) <- c("palb", "potilas", assumedpredictors_fat_epros$Name)
rows_with_missing_palb <- normalized_input_df[is.na(normalized_input_df$palb),]

```

```{r, palb_imputation2, echo=TRUE, eval=FALSE}
for (i in 1:nrow(rows_with_missing_palb)) {
  
  datarow <- rows_with_missing_palb[i,]
  
  # get model for patient in this datarow 
  
  personal_model_dir <- paste0("data_imputation_models/",datarow$potilas)
  print(paste0("Reading personal data imputation model '", personal_model_dir, 
               "/personal_graph.graphml'"))
  
  personal_graph <- read.graph(paste0(personal_model_dir, "/personal_graph.graphml"), "graphml")
  
  # and use the data in this row to predict palb
  evidence <- rows_with_missing_palb[i, assumedpredictors_fat_epros$Name]

  posterior_prediction <- mebn.personal_prediction(reaction_graph = personal_graph, 
                                 graph_dir = personal_model_dir,
                                 evidence = evidence,
                                 stan_model_file = "diet/posterior_prediction.stan")
  
  # personal_predictions contains predictions multivariate predictions
  # P-K and fP-Pi are known, but P-Alb is missing
  
  posterior <- rstan::extract(posterior_prediction, par= "posterior[3]")
  
  # use predicted posterior mean for missing P-Alb value
  # - this is i:th NA value in dialysis
  predicted_palb <- mean(posterior$`posterior[3]`)
  
  # rows in 'rows_with_missing_palb' and 'original_palb' are in same order
  # so we can imput in the NA values at 'original_palb'
  
  for (m in 1:imput_length)
  {
      if (is.na(original_palb[m]))
      {
          original_palb[m] <- predicted_palb
          break;
      }
  }
}

dialysis2 <- dialysis
dialysis2$palb <- original_palb
saveRDS(dialysis2, "data/DIALYSIS_imputed_palb.rds")
```

```{r, palb_imputation3, echo=FALSE, eval=FALSE}
write.csv2(dialysis2, file="data/DIALYSIS_imputed_palb.csv", row.names = FALSE, sep = ";")
```

With these data imputations we can estimate a cross-covariance model with three responses

```{r graph_with_gamma_qr_mv3_full_prep, echo=TRUE}
pk_fppi_palb_targets <- datadesc_fat_epros[datadesc_fat_epros$Name %in% c('pk','fppi','palb'),]

initial_graph <- mebn.fully_connected_bipartite_graph(datadesc_fat_epros)
```

```{r graph_with_gamma_qr_mv3_full, echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
dialysis_imputed <- readRDS("data/DIALYSIS_imputed_palb.rds")
no_holdout <- rep(0, nrow(dialysis_imputed))

dialdiet_gamma_mv3 <- mebn.bipartite_multivariate(reaction_graph = initial_graph, 
                                   inputdata = dialysis_imputed,
                                   targetdata = no_holdout,
                                   predictor_columns = assumedpredictors_fat_epros, 
                                   assumed_targets = pk_fppi_palb_targets, 
                                   group_column = "potilas",
                                   local_estimation = mebn.multivariate_sampling,
                                   local_model_cache = 
                                     "models/BLMM_gamma_qr_multivariate3/imputed_palb", 
                                   stan_model_file = "mebn/v2/BLMM_gamma_qr_mv_cv.stan",
                                   normalize_values = TRUE)

write.graph(dialdiet_gamma_mv3, 
            "graphs/dialysis_gamma_multivariate3_imputed.graphml", "graphml")
```

Finally, we consider also the effect of dialysis treatment type as new level of grouping in data. This allows us to estimate average effects of nutrition for each treatment type and then also personal effects within those treatments.

```{r graph_with_gamma_qr_mvcross_two_levels, echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
pk_fppi_palb_targets <- datadesc_fat_epros[datadesc_fat_epros$Name %in% c('pk','fppi','palb'),]
no_holdout <- rep(0, nrow(dialysis_imputed))

initial_graph <- mebn.fully_connected_bipartite_graph(datadesc_fat_epros)

dialysis_imputed <- readRDS("data/DIALYSIS_imputed_palb.rds")

# add the dialysis treatment type as a grouping factor
dialysis_imputed$hoitoryhma <- as.factor(dialysis_imputed$hoitomuoto)

# and sort the data by treatment/patient/observation
dialysis_imputed <- dialysis_imputed[order(dialysis_imputed$hoitoryhma, 
                                           dialysis_imputed$potilas, dialysis_imputed$havainto),]

dialdiet_gamma_mv3_two_level <- mebn.bipartite_two_level_multivariate(
                                   reaction_graph = initial_graph, 
                                   inputdata = dialysis_imputed,
                                   targetdata = no_holdout,
                                   predictor_columns = assumedpredictors_fat_epros, 
                                   assumed_targets = pk_fppi_palb_targets, 
                                   group_column = "hoitoryhma",
                                   subject_column = "potilas",
                                   local_estimation = mebn.two_level_multivariate_sampling,
                                   local_model_cache = 
                                     "models/BLMM_gamma_mv_cross/two_levels", 
                                   stan_model_file = 
                                     "mebn/v2/BLMM_gamma_two_level_grouping.stan",
                                   normalize_values = TRUE)

write.graph(dialdiet_gamma_mv3_two_level, 
            "graphs/dialysis_gamma_two_level_grouping.graphml", "graphml")
```

# Model analysis

Here we evaluate the model's ability to repeat the measured observations with normalized root mean square error (NRMSE) and compare how adding the dialysis treatment as an explicit hierarchical layer in the model affects this error. We also asses the model's performance with visual posterior predictive check by comparing samples from the model with measured values. This would give away possible bias in the model.


```{r extract_personal_generative_models, eval=FALSE, echo=FALSE}
source("mebn/v2/MEBNv2.r")

# Extract generative models for patients with data from personal grouping

# - latent parameters from this model
latent_parameter_modeldir <- "models/BLMM_gamma_mv_cross/two_levels"
#latent_parameter_modeldir <- "models/BLMM_gamma_qr_multivariate3/imputed_palb/"
#latent_parameter_modeldir <- "models/BLMM_gamma_separate/"

# - use these predictors and targets
datadesc <- datadesc_fat_epros
target_variables <- datadesc_fat_epros[datadesc_fat_epros$Name %in% c('pk','fppi','palb'),]
#target_variables <- datadesc_fat_epros[datadesc_fat_epros$Name %in% c('pk','fppi'),]
assumedpredictors <- assumedpredictors_fat_epros

# - output personal generative models (graphml + rv samples) in this dir
graph_dir <- "graphs/mv3_cross_two_levels/"
#graph_dir <- "graphs/mv3_cross/"
#graph_dir <- "graphs/separate_pk_fppi/"

# Loop through patients and produce the models

dialysis <- readRDS("data/DIALYSIS_imputed_palb.rds")

# add the dialysis treatment type as a grouping factor
dialysis$hoitoryhma <- as.factor(dialysis$hoitomuoto)

patients <- levels(dialysis$potilas)

#for (person_id in 1:length(patients)) {

for (person_id in 1:37) {
  
  # - initial graph structure
  initial_graph <- mebn.fully_connected_bipartite_graph(datadesc)
  
  # - pick the estimated latent variables for all the persons
  local_distributions <- target_variables
  local_distributions$modelcache <- latent_parameter_modeldir
  
  # - get personal data, normalized and original
  
  # - statistics for the nutrition levels are calculated from normalized data (as it was in the likelihood estimation)
  predictors <- nrow(assumedpredictors)
  normalized_input <- sapply(1:predictors, mebn.scale_gaussians, data = dialysis, datadesc = assumedpredictors, log_transform_ln = FALSE)
  normalized_input_df <- as.data.frame(normalized_input)
  
  # - pick rows for the selected person
  subject_code <- levels(dialysis$potilas)[person_id]
  personal_data_df <- cbind(dialysis$potilas, normalized_input_df)
  personal_data_df <- personal_data_df[personal_data_df$`dialysis$potilas` == subject_code,]
  personal_data <- as.matrix(subset(personal_data_df, select = -c(`dialysis$potilas`)))
  
  # - store also these original stats in graph
  personal_data_org <- subset(dialysis[dialysis$potilas == subject_code,], select = as.vector(assumedpredictors$Name))
  personal_concentrations_org <- subset(dialysis[dialysis$potilas == subject_code,], select = as.vector(assumedtargets$Name))
  
  personal_model_dir <- paste0(graph_dir, person_id)
  
  # Generate a personal graph in directory
  
  group_id <- as.numeric(unique(dialysis[dialysis$potilas == subject_code,]$hoitoryhma))

  # one-level, univariate models
  #personal_graph <- mebn.extract_personal_graph(person_id, initial_graph, personal_model_dir, assumedpredictors, target_variables, latent_parameter_modeldir, personal_data, personal_data_org)

  # one-level model
  #personal_graph <- mebn.extract_personal_graph_from_mv(person_id, initial_graph, personal_model_dir, assumedpredictors, assumedtargets, latent_parameter_modeldir, personal_data, personal_data_org, personal_concentrations_org, datadesc)

  # two-level model
  personal_graph <- mebn.extract_multilevel_graph(person_id, group_id, initial_graph, personal_model_dir, assumedpredictors, assumedtargets, latent_parameter_modeldir, personal_data, personal_data_org, personal_concentrations_org, datadesc)
  
}

```


```{r simulated_MSE, eval=FALSE, echo=FALSE}
library(igraph)
source("mebn/v2/MEBNv2.r")

potassium_org_mean <- mean(dialysis$kalium) # 2755
potassium_org_sd <- sd(dialysis$kalium) # 949
  
phosphorous_org_mean <- mean(dialysis$fosfori) # 2755
phosphorous_org_sd <- sd(dialysis$fosfori) # 949

pk_mean <- mean(dialysis$pk)
fppi_mean <- mean(dialysis$fppi)
palb_mean <- mean(dialysis$palb, na.rm = TRUE)

mse_summary <- data.frame(matrix(ncol = 5, nrow = 0), row.names = NULL)

#model_dir_candidates <- c("mv3_cross_two_levels", "mv3_cross")
model_dir_candidates <- c("separate_pk_fppi")

for (model_dir in model_dir_candidates) {

  graph_dir <- paste0("graphs/",model_dir,"/")

  for (person_id in 1:37) {
  
    print(paste0("Simulating patient ", person_id))
  
    subject_code <- levels(dialysis$potilas)[person_id]
    treatment <- unique(dialysis[dialysis$potilas == subject_code,]$hoitomuoto)
    
    # Normal ranges of concentration are personalized
    personal_info <- head(dialysis[dialysis$potilas == subject_code,],1)
    recommeded_concentrations <- mebn.get_personal_target_guidelines(personal_info, patient_in_dialysis = TRUE)
    
    # Omit P-Alb
    recommeded_concentrations$lower_limits <- recommeded_concentrations$lower_limits[1:2]
    recommeded_concentrations$upper_limits <- recommeded_concentrations$upper_limits[1:2]
    
    personal_model_dir <- paste0(graph_dir,person_id)
    personal_graph <- read.graph(paste0(personal_model_dir, "/personal_graph.graphml"), "graphml")
    
    # Sample diet proposals and matching concentration predictions
    
    # Omit P-Alb
    V(personal_graph)["palb"]$type <- 300
    
    repeat_model <- mebn.Query(reaction_graph = personal_graph,
                           graph_dir = personal_model_dir,
                           queried_nodes = c("kalium","fosfori"),
                           query = recommeded_concentrations,
                           proposal_limits = c(((0 - potassium_org_mean) / potassium_org_sd),
                                             ((5800 - potassium_org_mean) / potassium_org_sd),
                                             ((0 - phosphorous_org_mean) / phosphorous_org_sd),
                                             ((2550 - phosphorous_org_mean) / phosphorous_org_sd)),                         
                           stan_model_file = "diet/multivariate_intake_recommendation.stan",
                           conc_lower_limits = as.vector(recommeded_concentrations$lower_limits),
                           conc_upper_limits = as.vector(recommeded_concentrations$upper_limits),
                           beta_point_est = "mean",
                           param_point_est = "mean",
                           X_point_est = "mean",
                           posterior_samples = 1,
                           X_sd_coef = 1,
                           repeat_only = 1)
  
    repeats <- rstan::extract(repeat_model)
   
    pk <- repeats$concentration[,,1]
    fppi <- repeats$concentration[,,2]
    palb <- repeats$concentration[,,3]
     
    true_concs <- dialysis %>%
         filter(potilas == subject_code) %>%
         select(havainto, pk, fppi, palb)
    
    personal_mse <- cbind(subject_code,
                              model_dir,
                              mebn.NRMSE(mean(pk), mean(true_concs$pk), pk_mean),
                              mebn.NRMSE(mean(fppi), mean(true_concs$fppi), fppi_mean),
                              mebn.NRMSE(mean(palb), mean(true_concs$palb), palb_mean))

    # personal_mse <- cbind(subject_code,
    #                           model_dir,
    #                           mebn.NRMSE(mean(pk), mean(true_concs$pk), pk_mean),
    #                           mebn.NRMSE(mean(fppi), mean(true_concs$fppi), fppi_mean))

    mse_summary <- rbind(mse_summary, personal_mse)
  }
}

colnames(mse_summary) <- c("subject", "reaction_model", "pk_nrmse", "fppi_nrmse", "palb_nrmse")
#colnames(mse_summary) <- c("subject", "reaction_model", "pk_nrmse", "fppi_nrmse")

saveRDS(mse_summary, "model_evaluation/NRMSE_comparison_mv_models.rds")
#saveRDS(mse_summary, "model_evaluation/NRMSE_comparison_uni_model.rds")

write.csv2(mse_summary, file="model_evaluation/NRMSE_comparison_mv_models.csv", row.names = FALSE)
#write.csv2(mse_summary, file="model_evaluation/NRMSE_comparison_uni_model.csv", row.names = FALSE)

```

```{r MSE_table, cache=TRUE, echo=FALSE}

uni_mse_summary <- readRDS("model_evaluation/NRMSE_comparison_uni_model.rds")
mse_summary <- readRDS("model_evaluation/NRMSE_comparison_mv_models.rds")

uni_pk_mse <- mean(as.numeric(uni_mse_summary$pk_nrmse), na.rm=TRUE)
uni_fppi_mse <- mean(as.numeric(uni_mse_summary$fppi_nrmse), na.rm=TRUE)

mv3c2_pk_mse <- mean(as.numeric(mse_summary[mse_summary$reaction_model == "mv3_cross_two_levels",]$pk_nrmse)) # 0.004475995
mv3c2_fppi_mse <- mean(as.numeric(mse_summary[mse_summary$reaction_model == "mv3_cross_two_levels",]$fppi_nrmse)) # 0.006727212
mv3c2_palb_mse <- mean(as.numeric(mse_summary[mse_summary$reaction_model == "mv3_cross_two_levels",]$palb_nrmse), na.rm=TRUE) # 0.001819183

mv3c1_pk_mse <- mean(as.numeric(mse_summary[mse_summary$reaction_model == "mv3_cross",]$pk_nrmse), na.rm=TRUE) # 0.03623973
mv3c1_fppi_mse <- mean(as.numeric(mse_summary[mse_summary$reaction_model == "mv3_cross",]$fppi_nrmse), na.rm=TRUE) # 0.1298725
mv3c1_palb_mse <- mean(as.numeric(mse_summary[mse_summary$reaction_model == "mv3_cross",]$palb_nrmse), na.rm=TRUE) # 0.05903203

unic1 <- cbind("separate_pk_fppi", uni_pk_mse, uni_fppi_mse, 0, mean(uni_pk_mse, uni_fppi_mse))
mv3c1 <- cbind("mv3_cross_single_level", mv3c1_pk_mse, mv3c1_fppi_mse, mv3c1_palb_mse, mean(mv3c1_pk_mse, mv3c1_fppi_mse, mv3c1_palb_mse))
mv3c2 <- cbind("mv3_cross_two_levels", mv3c2_pk_mse, mv3c2_fppi_mse, mv3c2_palb_mse, mean(mv3c2_pk_mse, mv3c2_fppi_mse, mv3c2_palb_mse))

mse_table <- rbind(mv3c2, mv3c1, unic1)
mse_table[,2:5] <- round(as.numeric(mse_table[,2:5]),3)

kable(mse_table, "latex", escape = T, booktabs = T, row.names = FALSE, digits = 3,align="lrrrr",
  caption = "Comparison of model candidates. Personal graphical models with different reaction model candidates are evaluated with NRMSE for each modeled concentration and average model error.",
  col.names = c("Reaction model", "P-K", "fP-Pi", "P-Alb", "average error")) %>%
  add_header_above(c(" " = 1, "NRMSE" = 4), escape = F) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  kable_styling(latex_options = c("basic", "condensed"), full_width = FALSE) %>%
  row_spec(0,bold=TRUE) %>% save_kable(file = "tables/model_nrmse_comparison.pdf", keep_tex = TRUE)

```

```{r dialysis_gamma_mv_cross_2levels_ppc, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE, cache=TRUE, fig.cap="Posterior predictive check of the model where three concentrations and their parameters are stacked into one univariate model for estimating cross-model correlations."}
#source("mebn/v3/MEBNv3.r")
dialysis_imputed <- readRDS("data/DIALYSIS_imputed_palb.rds")
pk_fppi_palb_targets <- datadesc_fat_epros[datadesc_fat_epros$Name %in% c('pk','fppi','palb'),]

pk_fppi_palb_targets[pk_fppi_palb_targets$Name == "pk",]$ScaleMax <- 7.5
pk_fppi_palb_targets[pk_fppi_palb_targets$Name == "fppi",]$ScaleMax <- 5.0
pk_fppi_palb_targets[pk_fppi_palb_targets$Name == "palb",]$ScaleMax <- 50
pk_fppi_palb_targets

p <- mebn.multivariate_dens_overlays("models/BLMM_gamma_mv_cross/two_levels", pk_fppi_palb_targets, dialysis_imputed) 
ggsave("figures/ppc_mv3_cross.pdf", plot = p, width = 4, height = 4)
p
```

```{r, eval=FALSE, typical_effects_figure, fig.cap="This visualization shows mean of posterior for typical coefficients as weight of the connection between blood concentrations and nutrients at diet affecting it. The figure is plotted with iGraph package for R language (v 1.2.6, https://igraph.org/r).", fig.height = 4, fig.width=5, echo=FALSE, fig.align = "left", message=FALSE, warning=FALSE, cache=FALSE}

require(igraph)
source("mebn/v3/MEBNv3.r")
dialysis_mv3cross_treatment <- read_graph("graphs/dialysis_gamma_two_level_grouping.graphml", "graphml")

graph_layout <- mebn.plot_typical_effects(dialysis_mv3cross_treatment, 20, graph_layout = NULL)
```

Then we look into the estimated correlation matrices of effects. The model estimates how much the effects of nutrients vary between different types of dialysis treatment and further how the effects vary personally within a treatment. These correlation plots show how the effects are correlated. 

```{r, subject_effect_correlations, eval=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
library(rstan)
library(ggcorrplot)

#target_blmm <- mebn.get_localfit("pk_fppi_palb", "models/BLMM_gamma_mv_cross")
target_blmm <- mebn.get_localfit("pk_fppi_palb", "models/BLMM_gamma_mv_cross/two_levels")

C_extract <- rstan::extract(target_blmm, pars = c("C_s"))

C_cross_draws <- C_extract$C[,,]
C_cross <- apply(C_cross_draws, 2:3, mean)

# remove variance from diagonal
diag(C_cross) <- 0

C_cross.df <- as.data.frame(C_cross)

colnames(C_cross.df) <- c("P-K level", paste(as.vector(assumedpredictors_fat_epros$Description), " -> P-K", sep = ""), "fP-Pi level", paste(as.vector(assumedpredictors_fat_epros$Description), " -> fP-Pi", sep = ""), "P-Alb level", paste(as.vector(assumedpredictors_fat_epros$Description), " -> P-Alb", sep = ""))
rownames(C_cross.df) <- c("P-K level", paste(as.vector(assumedpredictors_fat_epros$Description), " -> P-K", sep = ""), "fP-Pi level", paste(as.vector(assumedpredictors_fat_epros$Description), " -> fP-Pi", sep = ""), "P-Alb level", paste(as.vector(assumedpredictors_fat_epros$Description), "-> P-Alb", sep = ""))

saveRDS(C_cross.df, "model_evaluation/subject_effect_correlations_df.rds")

# Custom corrplot
# library(reshape2)   
# 
# C_cross.df
# 
# ggplot(melt(C_cross.df), aes(Var1, Var2, fill=value)) +
#   geom_tile(height=0.8, width=0.8) +
#   scale_fill_gradient2(low="blue", mid="white", high="red") +
#   theme_minimal() +
#   coord_equal() +
#   labs(x="",y="",fill="Corr") +
#   theme(axis.text.x=element_text(size=13, angle=45, vjust=1, hjust=1, 
#                                  margin=margin(-3,0,0,0)),
#         axis.text.y=element_text(size=13, margin=margin(0,-3,0,0)),
#         panel.grid.major=element_blank()) 

```

```{r, treatment_effect_correlations, eval=FALSE, echo=FALSE}
library(rstan)
library(ggcorrplot)

#target_blmm <- mebn.get_localfit("pk_fppi_palb", "models/BLMM_gamma_mv_cross")
target_blmm <- mebn.get_localfit("pk_fppi_palb", "models/BLMM_gamma_mv_cross/two_levels")

C_extract <- rstan::extract(target_blmm, pars = c("C_g"))

C_cross_draws <- C_extract$C[,,]
C_cross <- apply(C_cross_draws, 2:3, mean)

# remove variance from diagonal
diag(C_cross) <- 0

C_cross.df <- as.data.frame(C_cross)

colnames(C_cross.df) <- c("P-K level", paste(as.vector(assumedpredictors_fat_epros$Description), " -> P-K", sep = ""), "fP-Pi level", paste(as.vector(assumedpredictors_fat_epros$Description), " -> fP-Pi", sep = ""), "P-Alb level", paste(as.vector(assumedpredictors_fat_epros$Description), " -> P-Alb", sep = ""))
rownames(C_cross.df) <- c("P-K level", paste(as.vector(assumedpredictors_fat_epros$Description), " -> P-K", sep = ""), "fP-Pi level", paste(as.vector(assumedpredictors_fat_epros$Description), " -> fP-Pi", sep = ""), "P-Alb level", paste(as.vector(assumedpredictors_fat_epros$Description), "-> P-Alb", sep = ""))

saveRDS(C_cross.df, "model_evaluation/dialysis_effect_correlations_df.rds")

```

```{r, treatment_effect_correlation_plot, fig.width=14, fig.out="100%", fig.align="left", eval=TRUE, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
library(ggcorrplot)

C_cross.df <- readRDS("model_evaluation/dialysis_effect_correlations_df.rds")

p <- nrow(assumedpredictors_fat_epros) + 1

cplot <- ggcorrplot(C_cross.df, hc.order = FALSE, lab = FALSE, type="lower") +
  scale_fill_gradient2(limit = c(min(C_cross.df),max(C_cross.df)), low = "blue", high = "red", mid = "white", midpoint = 0) +
  ggplot2::labs(x = NULL, y = NULL) +  
  ggplot2::theme(
    plot.title = element_text(size = 10),
    axis.title.x = element_text(angle = 0, color = 'grey20'),
    axis.title.y = element_text(angle = 90, color = 'grey20'),
    axis.text.x = element_text(size = 8),
    axis.text.y = element_text(size = 8),
    legend.text = element_text(size = 8),
    legend.position = "bottom",
    legend.title = element_blank(),
    plot.margin = margin(t = 0,  # Top margin
                         r = 0,  # Right margin
                         b = 0,  # Bottom margin
                         l = 0) # Left margin
 ) +
ggplot2::geom_vline(xintercept = c(p,2*p), linetype="solid", color = "grey", size=1) +
ggplot2::geom_hline(yintercept = c(p,2*p), linetype="solid", color = "grey", size=1)

ggsave(paste0("figures/treatment_level_reaction_correlations.pdf"), plot = cplot, width = 15, height = 15, scale = 0.7)

```
```{r, subject_effect_correlation_plot, fig.width=14, fig.out="100%", fig.align="left", eval=TRUE, cache=FALSE, echo=FALSE}
library(ggcorrplot)

C_cross.df <- readRDS("model_evaluation/subject_effect_correlations_df.rds")

p <- nrow(assumedpredictors_fat_epros) + 1

cplot <- ggcorrplot(C_cross.df, hc.order = FALSE, lab = FALSE, type="lower") +
  scale_fill_gradient2(limit = c(min(C_cross.df),max(C_cross.df)), low = "blue", high = "red", mid = "white", midpoint = 0) +
  ggplot2::labs(x = NULL, y = NULL) +  
  ggplot2::theme(
    plot.title = element_text(size = 10),
    axis.title.x = element_text(angle = 0, color = 'grey20'),
    axis.title.y = element_text(angle = 90, color = 'grey20'),
    axis.text.x = element_text(size = 8),
    axis.text.y = element_text(size = 8),
    legend.text = element_text(size = 8),
    legend.position = "bottom",
    legend.title = element_blank(),
    plot.margin = margin(t = 0,  # Top margin
                         r = 0,  # Right margin
                         b = 0,  # Bottom margin
                         l = 0) # Left margin
 ) +
ggplot2::geom_vline(xintercept = c(p,2*p), linetype="solid", color = "grey", size=1) +
ggplot2::geom_hline(yintercept = c(p,2*p), linetype="solid", color = "grey", size=1)

ggsave(paste0("figures/subject_level_reaction_correlations.pdf"), plot = cplot, width = 15, height = 15, scale = 0.7)

```

```{r, include=TRUE, fig.align="center", fig.cap=c("Correlation plot of matrix $C_g$ that includes the estimated within-model and cross-model correlations in level of dialysis treatment"), echo=FALSE}
knitr::include_graphics("figures/treatment_level_reaction_correlations.pdf")
```

```{r, include=TRUE, fig.align="center", fig.cap=c("Correlation plot of matrix $C_b$ that includes the estimated within-model and cross-model correlations of personal effects"), echo=FALSE}
knitr::include_graphics("figures/subject_level_reaction_correlations.pdf")
```

```{r, top_treatment_correlations, eval=TRUE, cache=FALSE, echo=FALSE}
C_cross_g.df <- readRDS("model_evaluation/dialysis_effect_correlations_df.rds")

effect_cor.df <- data.frame(matrix(ncol = 3, nrow = 0), row.names = NULL)

for (x in 1:dim(C_cross_g.df)[1])
{
  for (y in 1:dim(C_cross_g.df)[2])
  {
    eff <- C_cross_g.df[x,y]
    from <- colnames(C_cross_g.df)[x]
    to <- rownames(C_cross_g.df)[y]
    
    effect_cor.df <- rbind(effect_cor.df, c(from,to,round(eff,3),abs(eff)))
  }
}

colnames(effect_cor.df) <- c("effect1", "effect2", "corr", "abs_effect")

effect_cor.df$corr <- as.numeric(effect_cor.df$corr)
effect_cor.df$abs_effect <- as.numeric(effect_cor.df$abs_effect)

# fix effect names for table
effect_cor.df$effect1 <- unlist(lapply(strsplit(effect_cor.df$effect1, ":"), function(x) ifelse(is.na(x[2]), x[1], paste0(x[2]," -> ",x[1]))))
effect_cor.df$effect2 <- unlist(lapply(strsplit(effect_cor.df$effect2, ":"), function(x) ifelse(is.na(x[2]), x[1], paste0(x[2]," -> ",x[1]))))

top_correlations <- effect_cor.df %>% 
  top_n(60, abs_effect) %>%
  select("effect1","effect2","corr") %>%
  arrange(-corr) %>%
  slice(which(row_number() %% 2 == 1))

#kable(top_correlations, "latex", escape = T, booktabs = T, row.names = FALSE, digits = 3, col.names = c("Effect 1", "Effect 2", "Correlation"), caption = "Table shows 30 highest positive or negative correlations between effects of dialysis treatment type.") %>%
#    kable_styling(latex_options = "HOLD_position") %>%
#    kable_styling(latex_options = c("basic", "condensed"), full_width = FALSE) %>%
#    add_header_above(c("Effects of nutrients in dialysis treatment level" = 2, " " = 1), escape = F) %>%
#    row_spec(0,bold=TRUE)

top_pot_correlations <- effect_cor.df %>%
  filter(substr(effect2,1,9) == "Potassium") %>%
  top_n(40, abs_effect) %>%
  select("effect1","effect2","corr") %>%
  arrange(-corr) %>%
  slice(which(row_number() %% 2 == 1))

top_pho_correlations <- effect_cor.df %>%
  filter(substr(effect2,1,11) == "Phosphorous") %>%
  top_n(40, abs_effect) %>%
  select("effect1","effect2","corr") %>%
  arrange(-corr) %>%
  slice(which(row_number() %% 2 == 1))

kable(rbind(top_pot_correlations, top_pho_correlations), "latex", escape = T, booktabs = T, row.names = FALSE, digits = 3, col.names = c("Effect 1", "Effect 2", "Correlation"), caption = "Table shows 40 highest positive or negative correlations between treatment effects of potassium and phosphorous with other treatment effects. This structure of correlations is used in estimating the personal effects based on personal intake and matching concentrations.") %>%
    kable_styling(latex_options = "HOLD_position") %>%
    kable_styling(latex_options = c("basic", "condensed"), full_width = FALSE) %>%
    add_header_above(c("Treatment-level effects of nutrients" = 2, " " = 1), escape = F) %>%
    row_spec(0,bold=TRUE)
  
  #    save_kable(file = "tables/treatment_pot_and_pho_effect_correlations.pdf", keep_tex = TRUE)

```


```{r, top_subject_correlations, eval=TRUE, cache=FALSE, echo=FALSE}
C_cross_s.df <- readRDS("model_evaluation/subject_effect_correlations_df.rds")

effect_cor.df <- data.frame(matrix(ncol = 3, nrow = 0), row.names = NULL)

for (x in 1:dim(C_cross_s.df)[1])
{
  for (y in 1:dim(C_cross_s.df)[2])
  {
    eff <- C_cross_s.df[x,y]
    from <- colnames(C_cross_s.df)[x]
    to <- rownames(C_cross_s.df)[y]
    
    effect_cor.df <- rbind(effect_cor.df, c(from,to,round(eff,3),abs(eff)))
  }
}

colnames(effect_cor.df) <- c("effect1", "effect2", "corr", "abs_effect")

effect_cor.df$corr <- as.numeric(effect_cor.df$corr)
effect_cor.df$abs_effect <- as.numeric(effect_cor.df$abs_effect)

# fix effect names for table
effect_cor.df$effect1 <- unlist(lapply(strsplit(effect_cor.df$effect1, ":"), function(x) ifelse(is.na(x[2]), x[1], paste0(x[2]," -> ",x[1]))))
effect_cor.df$effect2 <- unlist(lapply(strsplit(effect_cor.df$effect2, ":"), function(x) ifelse(is.na(x[2]), x[1], paste0(x[2]," -> ",x[1]))))

top_correlations <- effect_cor.df %>% 
  top_n(60, abs_effect) %>%
  select("effect1","effect2","corr") %>%
  arrange(-corr) %>%
  slice(which(row_number() %% 2 == 1))

#kable(top_correlations, "latex", escape = T, booktabs = T, row.names = FALSE, digits = 3, col.names = c("Effect 1", "Effect 2", "Correlation"), caption = "Table shows 30 highest positive or negative correlations between personal effects of nutrients or other considered predictors. This structure of correlations is used in estimating the personal effects based on personal intake and matching concentrations.") %>%
#    kable_styling(latex_options = "HOLD_position") %>%
#    kable_styling(latex_options = c("basic", "condensed"), full_width = FALSE) %>%
#    add_header_above(c("Personal effects of nutrients" = 2, " " = 1), escape = F) %>%
#    row_spec(0,bold=TRUE)


top_pot_correlations <- effect_cor.df %>%
  filter(substr(effect2,1,9) == "Potassium") %>%
  top_n(40, abs_effect) %>%
  select("effect1","effect2","corr") %>%
  arrange(-corr) %>%
  slice(which(row_number() %% 2 == 1))

top_pho_correlations <- effect_cor.df %>%
  filter(substr(effect2,1,11) == "Phosphorous") %>%
  top_n(40, abs_effect) %>%
  select("effect1","effect2","corr") %>%
  arrange(-corr) %>%
  slice(which(row_number() %% 2 == 1))

kable(rbind(top_pot_correlations, top_pho_correlations), "latex", escape = T, booktabs = T, row.names = FALSE, digits = 3, col.names = c("Effect 1", "Effect 2", "Correlation"), caption = "Table shows 40 highest positive or negative correlations between personal effects of potassium and phosphorous with other personal effects. This structure of correlations is used in estimating the personal effects based on personal intake and matching concentrations.") %>%
    kable_styling(latex_options = "HOLD_position") %>%
    kable_styling(latex_options = c("basic", "condensed"), full_width = FALSE) %>%
    add_header_above(c("Personal effects of nutrients" = 2, " " = 1), escape = F) %>%
    row_spec(0,bold=TRUE)

#    save_kable(file = "tables/subject_pot_and_pho_effect_correlations.pdf", keep_tex = TRUE)

```

```{r build_two_level_effect_summary, echo=FALSE, eval=FALSE}

library(igraph)
graph_dir <- "graphs/mv3_cross_two_levels/"

# - collect personal statistics in this data frame
effect_summary <- data.frame(matrix(ncol = 17, nrow = 0), row.names = NULL)

for (person_id in 1:37) {
    
  personal_model_dir <- paste0(graph_dir,person_id)
  personal_graph <- read.graph(paste0(personal_model_dir, "/personal_graph.graphml"), "graphml")
  
  subject_code <- levels(dialysis$potilas)[person_id]
  treatment <- unique(dialysis[dialysis$potilas == subject_code,]$hoitomuoto)

  for (c in 1:dim(assumedtargets)[1])
  {
    target_column <- assumedtargets[c,]
    target_name <- as.vector(target_column$Name)
    
    predictor_names <- as.vector(assumedpredictors_fat_epros$Name)
    
    for (p in 1:dim(assumedpredictors_fat_epros)[1])
    {
      predictor_name <- predictor_names[p]
    
      effect_name <- paste0(predictor_name,"_",target_name)
        
      nodes <- V(personal_graph)
  
      personal_effects <- c(person_id,
                            effect_name,
                            nodes[paste0(effect_name)]$value,
                            nodes[paste0(effect_name)]$mode,
                            nodes[paste0(effect_name)]$value_lCI,
                            nodes[paste0(effect_name)]$value_uCI,
                            nodes[paste0("sigma_g_", effect_name)]$value,
                            nodes[paste0("sigma_g_", effect_name)]$value_lCI,
                            nodes[paste0("sigma_g_", effect_name)]$value_uCI,
                            nodes[paste0("sigma_b_", effect_name)]$value,
                            nodes[paste0("sigma_b_", effect_name)]$value_lCI,
                            nodes[paste0("sigma_b_", effect_name)]$value_uCI,
                            treatment,
                            nodes[paste0("personal_", effect_name)]$group,
                            nodes[paste0("personal_", effect_name)]$group_mode,
                            nodes[paste0("personal_", effect_name)]$group_lCI,
                            nodes[paste0("personal_", effect_name)]$group_uCI,
                            nodes[paste0("personal_", effect_name)]$value,
                            nodes[paste0("personal_", effect_name)]$mode,
                            nodes[paste0("personal_", effect_name)]$value_lCI,
                            nodes[paste0("personal_", effect_name)]$value_uCI)
  
  
      effect_summary <- rbind(effect_summary,personal_effects)
    }
  }

}

colnames(effect_summary) <- c("person_id", "effect", "general_effect", "general_effect_mode", "general_effect_lCI", "general_effect_uCI", "sigma_g", "sigma_g_lCI", "sigma_g_uCI", "sigma_b", "sigma_b_lCI", "sigma_b_uCI", "treatment_group", "group_effect", "group_effect_mode", "group_effect_lCI", "group_effect_uCI", "personal_effect", "personal_effect_mode","personal_effect_lCI","personal_effect_uCI")

saveRDS(effect_summary, "patient_summary/two_level_effect_summary.rds")

```

# Overview of nutritional effects

```{r effect_summary_table, eval=TRUE, cache=FALSE, echo=FALSE}
library(dplyr)

effect_summary <- readRDS("patient_summary/two_level_effect_summary.rds")

general_summary <- effect_summary %>% 
  arrange(effect) %>%
  select(effect, 
         general_effect, 
         general_effect_mode, 
         general_effect_lCI, 
         general_effect_uCI, 
         sigma_g, 
         sigma_g_lCI, 
         sigma_g_uCI, 
         sigma_b, 
         sigma_b_lCI, 
         sigma_b_uCI) %>%
  group_by(effect) %>%
  dplyr::summarise(
                general_effect = as.numeric(dplyr::first(general_effect)),
                general_effect_mode = as.numeric(dplyr::first(general_effect_mode)),
                general_effect_lCI = as.numeric(dplyr::first(general_effect_lCI)),
                general_effect_uCI = as.numeric(dplyr::first(general_effect_uCI)),
                sigma_g = as.numeric(dplyr::first(sigma_g)),
                sigma_g_lCI = as.numeric(dplyr::first(sigma_g_lCI)),
                sigma_g_uCI = as.numeric(dplyr::first(sigma_g_uCI)),
                sigma_b = as.numeric(dplyr::first(sigma_b)),
                sigma_b_lCI = as.numeric(dplyr::first(sigma_b_lCI)),
                sigma_b_uCI = as.numeric(dplyr::first(sigma_b_uCI))
                )

KHD_summary <- effect_summary %>% 
  arrange(effect) %>%
  select(effect, 
         treatment_group,
         group_effect,
         group_effect_mode,
         group_effect_lCI,
         group_effect_uCI,
         personal_effect,
         personal_effect_mode,
         personal_effect_lCI,
         personal_effect_uCI) %>%
  filter(treatment_group == "KHD") %>%
  group_by(effect) %>%
  dplyr::summarise(
                KHD_effect = as.numeric(dplyr::first(group_effect)),
                KHD_effect_mode = as.numeric(dplyr::first(group_effect_mode)),
                KHD_effect_lCI = as.numeric(dplyr::first(group_effect_lCI)),
                KHD_effect_uCI = as.numeric(dplyr::first(group_effect_uCI)),
                min_personal_KHD = min(as.numeric(personal_effect)),
                min_personal_mode_KHD = min(as.numeric(personal_effect_mode)),
                min_personal_KHD_lCI = as.numeric(personal_effect_lCI[which.min(as.numeric(personal_effect))]),
                min_personal_KHD_uCI = as.numeric(personal_effect_uCI[which.min(as.numeric(personal_effect))]),
                max_personal_KHD = max(as.numeric(personal_effect)),
                max_personal_mode_KHD = max(as.numeric(personal_effect_mode)),
                max_personal_KHD_lCI = as.numeric(personal_effect_lCI[which.max(as.numeric(personal_effect))]),
                max_personal_KHD_uCI = as.numeric(personal_effect_uCI[which.max(as.numeric(personal_effect))])
  )

OHD_summary <- effect_summary %>% 
  arrange(effect) %>%
  select(effect, 
         treatment_group,
         group_effect,
         group_effect_mode,
         group_effect_lCI,
         group_effect_uCI,
         personal_effect,
         personal_effect_mode,
         personal_effect_lCI,
         personal_effect_uCI) %>%
  filter(treatment_group == "OHD") %>%
  group_by(effect) %>%
  dplyr::summarise(
                OHD_effect = as.numeric(dplyr::first(group_effect)),
                OHD_effect_mode = as.numeric(dplyr::first(group_effect_mode)),
                OHD_effect_lCI = as.numeric(dplyr::first(group_effect_lCI)),
                OHD_effect_uCI = as.numeric(dplyr::first(group_effect_uCI)),
                min_personal_OHD = min(as.numeric(personal_effect)),
                min_personal_mode_OHD = min(as.numeric(personal_effect_mode)),
                min_personal_OHD_lCI = as.numeric(personal_effect_lCI[which.min(as.numeric(personal_effect))]),
                min_personal_OHD_uCI = as.numeric(personal_effect_uCI[which.min(as.numeric(personal_effect))]),
                max_personal_OHD = max(as.numeric(personal_effect)),
                max_personal_mode_OHD = max(as.numeric(personal_effect_mode)),
                max_personal_OHD_lCI = as.numeric(personal_effect_lCI[which.max(as.numeric(personal_effect))]),
                max_personal_OHD_uCI = as.numeric(personal_effect_uCI[which.max(as.numeric(personal_effect))])
  )

PD_summary <- effect_summary %>% 
  arrange(effect) %>%
  select(effect, 
         treatment_group,
         group_effect,
         group_effect_mode,
         group_effect_lCI,
         group_effect_uCI,
         personal_effect,
         personal_effect_mode,
         personal_effect_lCI,
         personal_effect_uCI) %>%
  filter(treatment_group == "PD") %>%
  group_by(effect) %>%
  dplyr::summarise(
                PD_effect = as.numeric(dplyr::first(group_effect)),
                PD_effect_mode = as.numeric(dplyr::first(group_effect_mode)),
                PD_effect_lCI = as.numeric(dplyr::first(group_effect_lCI)),
                PD_effect_uCI = as.numeric(dplyr::first(group_effect_uCI)),
                min_personal_PD = min(as.numeric(personal_effect)),
                min_personal_mode_PD = min(as.numeric(personal_effect_mode)),
                min_personal_PD_lCI = as.numeric(personal_effect_lCI[which.min(as.numeric(personal_effect))]),
                min_personal_PD_uCI = as.numeric(personal_effect_uCI[which.min(as.numeric(personal_effect))]),
                max_personal_PD = max(as.numeric(personal_effect)),
                max_personal_mode_PD = max(as.numeric(personal_effect_mode)),
                max_personal_PD_lCI = as.numeric(personal_effect_lCI[which.max(as.numeric(personal_effect))]),
                max_personal_PD_uCI = as.numeric(personal_effect_uCI[which.max(as.numeric(personal_effect))])
  )

# lookup dataframe for pretty effect names
all_effects <- expand.grid(predictor = assumedpredictors_fat_epros$Name, target = assumedtargets$Name)
all_descs <- expand.grid(predictor = assumedpredictors_fat_epros$Description, target = assumedtargets$Description)
effect_desc.df <- data.frame(effect <- paste0(all_effects$predictor,"_",all_effects$target), row.names = NULL)
effect_desc.df$desc <- paste0(effect <- paste0(all_descs$predictor," -> ",all_descs$target))
effect_desc.df$nutrient <- all_descs$predictor
effect_desc.df$concentration <- all_descs$target
colnames(effect_desc.df) <- c("effect", "effect_description", "nutrient", "concentration")

full_summary <- general_summary %>% 
  inner_join(KHD_summary, by = "effect") %>%
  inner_join(OHD_summary, by = "effect") %>%
  inner_join(PD_summary, by = "effect") %>%
  inner_join(effect_desc.df, by = "effect")

# new df for table
two_level_effects_table.df <- data.frame(effect <- full_summary$effect, row.names = NULL)

# escape %-sign
#two_level_effects_table.df$effect_description <- gsub("%"," ",full_summary$effect_description)
two_level_effects_table.df$effect_description <- full_summary$effect_description

two_level_effects_table.df$nutrient <- full_summary$nutrient
two_level_effects_table.df$concentration <- full_summary$concentration

two_level_effects_table.df$general_effect <- linebreak(paste0(format(round(full_summary$general_effect,2), nsmall=2),"\n$\\left[",format(round(full_summary$general_effect_lCI,2), nsmall=2),"; ",format(round(full_summary$general_effect_uCI,2), nsmall=2),"\\right]$"), align = "c")

# - adds mode
#two_level_effects_table.df$general_effect <- linebreak(paste0(format(round(full_summary$general_effect,2), nsmall=2),", ",format(round(full_summary$general_effect_mode,2), nsmall=2), "\n$\\left[",format(round(full_summary$general_effect_lCI,2), nsmall=2),"; ",format(round(full_summary$general_effect_uCI,2), nsmall=2),"\\right]$"), align = "c")

two_level_effects_table.df$sigma_g <- linebreak(paste0(format(round(full_summary$sigma_g,2), nsmall = 2),"\n$\\left[",format(round(full_summary$sigma_g_lCI,2), nsmall = 2),"; ",format(round(full_summary$sigma_g_uCI,2), nsmall = 2),"\\right]$"), align = "c")
two_level_effects_table.df$sigma_b <- linebreak(paste0(format(round(full_summary$sigma_b,2), nsmall = 2),"\n$\\left[",format(round(full_summary$sigma_b_lCI,2), nsmall = 2),"; ",format(round(full_summary$sigma_b_uCI,2), nsmall = 2),"\\right]$"), align = "c")

two_level_effects_table.df$KHD <- linebreak(paste0(format(round(full_summary$KHD_effect,2), nsmall = 2),"\n$\\left[",format(round(full_summary$KHD_effect_lCI,2), nsmall = 2),"; ",format(round(full_summary$KHD_effect_uCI,2), nsmall = 2),"\\right]$"), align = "c")
two_level_effects_table.df$KHD_min <- linebreak(paste0(format(round(full_summary$min_personal_KHD,2), nsmall = 2),"\n$\\left[",format(round(full_summary$min_personal_KHD_lCI,2), nsmall = 2),"; ",format(round(full_summary$min_personal_KHD_uCI,2), nsmall = 2),"\\right]$"), align = "c")
two_level_effects_table.df$KHD_max <- linebreak(paste0(format(round(full_summary$max_personal_KHD,2), nsmall = 2),"\n$\\left[",format(round(full_summary$max_personal_KHD_lCI,2), nsmall = 2),"; ",format(round(full_summary$max_personal_KHD_uCI,2), nsmall = 2),"\\right]$"), align = "c")

two_level_effects_table.df$OHD <- linebreak(paste0(format(round(full_summary$OHD_effect,2), nsmall = 2),"\n$\\left[",format(round(full_summary$OHD_effect_lCI,2), nsmall = 2),"; ",format(round(full_summary$OHD_effect_uCI,2), nsmall = 2),"\\right]$"), align = "c")
two_level_effects_table.df$OHD_min <- linebreak(paste0(format(round(full_summary$min_personal_OHD,2), nsmall = 2),"\n$\\left[",format(round(full_summary$min_personal_OHD_lCI,2), nsmall = 2),"; ",format(round(full_summary$min_personal_OHD_uCI,2), nsmall = 2),"\\right]$"), align = "c")
two_level_effects_table.df$OHD_max <- linebreak(paste0(format(round(full_summary$max_personal_OHD,2), nsmall = 2),"\n$\\left[",format(round(full_summary$max_personal_OHD_lCI,2), nsmall = 2),"; ",format(round(full_summary$max_personal_OHD_uCI,2), nsmall = 2),"\\right]$"), align = "c")

two_level_effects_table.df$PD <- linebreak(paste0(format(round(full_summary$PD_effect,2), nsmall = 2),"\n$\\left[",format(round(full_summary$PD_effect_lCI,2), nsmall = 2),"; ",format(round(full_summary$PD_effect_uCI,2), nsmall = 2),"\\right]$"), align = "c")
two_level_effects_table.df$PD_min <- linebreak(paste0(format(round(full_summary$min_personal_PD,2), nsmall = 2),"\n$\\left[",format(round(full_summary$min_personal_PD_lCI,2), nsmall = 2),"; ",format(round(full_summary$min_personal_PD_uCI,2), nsmall = 2),"\\right]$"), align = "c")
two_level_effects_table.df$PD_max <- linebreak(paste0(format(round(full_summary$max_personal_PD,2), nsmall = 2),"\n$\\left[",format(round(full_summary$max_personal_PD_lCI,2), nsmall = 2),"; ",format(round(full_summary$max_personal_PD_uCI,2), nsmall = 2),"\\right]$"), align = "c")

# for sorting
two_level_effects_table.df$sg <- full_summary$sigma_g
two_level_effects_table.df$sb <- full_summary$sigma_b
two_level_effects_table.df$abs_general <- abs(full_summary$general_effect)

# Sort table by general effect level and variation between treatments

two_level_effects_table.df <- two_level_effects_table.df[order(-two_level_effects_table.df$sg),] 

```

```{r pot_pho_effects_only, cache=FALSE, echo=FALSE, eval=FALSE}

# This table is used in the article

pot_pho_effects.df <- two_level_effects_table.df %>%
  filter(nutrient == "Potassium" | nutrient == "Phosphorous") %>%
  arrange(nutrient)

kable(pot_pho_effects.df[c(3:4,5,8:16)], "latex", escape = F, booktabs = T, row.names = FALSE, digits = 2,align="llccccccccccc",
  caption = "Effects of potassium and phosphorous between different dialysis treatments and between patients within the same treatment.",
  col.names = c("Nutrient", "Conc.", "General effect", "avg", "min", "max", "avg", "min", "max", "avg", "min", "max")) %>%
  column_spec(1, width = "7em") %>%
  column_spec(2, width = "4em") %>%
  column_spec(4, bold=TRUE) %>%
  column_spec(7, bold=TRUE) %>%
  column_spec(10, bold=TRUE) %>%
  add_header_above(c(" " = 3, "Home hemodialysis" = 3, "Hospital hemodialysis" = 3, "Peritoneal dialysis" = 3), escape = F) %>%
 kable_styling(latex_options="scale_down") %>%  
  kable_styling(latex_options = c("basic", "condensed"), full_width = FALSE) %>%
  row_spec(0,bold=TRUE) %>%
  save_kable(file = "tables/pot_and_pho_effects.pdf", keep_tex = TRUE)
```

Following tables include all the estimated nutrition effects in general, dialysis treatment and personal levels.  
```{r full_effects_table, cache=FALSE, echo=FALSE}
# kable(two_level_effects_table.df[2:14], "latex", booktabs = T, row.names = FALSE, digits = 2,align="lrrrrrrrrrrrr",
#   caption = "Variation of nutrition effects between different dialysis treatments and between patients within the same treatment.",
#   col.names = c("Effect", "General effect", "between treatments", "within treatments", "avg", "min", "max", "avg", "min", "max", "avg", "min", "max")) %>%
#   add_header_above(c(" " = 2, "Variation" = 2, "KHD" = 3, "OHD" = 3, "PD" = 3), escape = F) %>%
#  kable_styling(latex_options="scale_down") %>%  
#   kable_styling(latex_options = c("basic", "condensed"), full_width = TRUE) %>%
#   row_spec(0,bold=TRUE) %>%
#   save_kable(file = "tables/variation_of_nutrition_effects2.pdf", keep_tex = TRUE)

# variations removed 
#  caption = "",

kabletable1 <- two_level_effects_table.df[1:32,c(3:4,5,8:16)]
kabletable1$nutrient <- gsub("%", "\\\\%", kabletable1$nutrient)

table1 <- kable(kabletable1, "latex", escape = F, booktabs = T, row.names = FALSE, digits = 2,align="llccccccccccc",
  col.names = c("Nutrient", "Conc.", "General effect", "avg", "min", "max", "avg", "min", "max", "avg", "min", "max")) %>%
  column_spec(1, width = "7em") %>%
  column_spec(2, width = "4em") %>%
  column_spec(4, bold=TRUE) %>%
  column_spec(7, bold=TRUE) %>%
  column_spec(10, bold=TRUE) %>%
  add_header_above(c(" " = 3, "Home hemodialysis" = 3, "Hospital hemodialysis" = 3, "Peritoneal dialysis" = 3), escape = F) %>%
  row_spec(0,bold=TRUE) %>%
  kable_styling(latex_options = "scale_down") %>%
  kable_styling(latex_options = c("basic", "condensed"))

table1 %>%  save_kable(file = "tables/variation_of_nutrition_effects1.pdf", keep_tex = TRUE)

kabletable2 <- two_level_effects_table.df[33:66,c(3:4,5,8:16)]
kabletable2$nutrient <- gsub("%", "\\\\%", kabletable2$nutrient)

table2 <- kable(kabletable2, "latex", escape = F, booktabs = T, row.names = FALSE, digits = 2,align="llccccccccccc",
  col.names = c("Nutrient", "Conc.", "General effect", "avg", "min", "max", "avg", "min", "max", "avg", "min", "max")) %>%
  column_spec(1, width = "7em") %>%
  column_spec(2, width = "4em") %>%
  column_spec(4, bold=TRUE) %>%
  column_spec(7, bold=TRUE) %>%
  column_spec(10, bold=TRUE) %>%
  add_header_above(c(" " = 3, "Home hemodialysis" = 3, "Hospital hemodialysis" = 3, "Peritoneal dialysis" = 3), escape = F) %>%
  row_spec(0,bold=TRUE) %>%
  kable_styling(latex_options = "scale_down") %>%
  kable_styling(latex_options = c("basic", "condensed")) 

table2 %>% save_kable(file = "tables/variation_of_nutrition_effects2.pdf", keep_tex = TRUE)

```
Supplementary Table S4: Nutrition effect magnitudes from nutrients and other modeled features $(j = 1,\dots,22)$ to blood concentrations $(i = 1,\dots,3)$ for analyzed patients $(p = 1,\dots,37)$ in all three additive levels of the model. General effects ($\hat{\beta}_{ij}$) are shown to vary between patients in home hemodialysis, hospital hemodialysis, and peritoneal dialysis. First column of each dialysis type (avg) shows typical effect of the treatment ($\hat{\beta}_{ij} + \hat{g}_{ijk}, k = 1,\dots,3$) that can further vary personally. Minimum and maximum of these personal effects are shown within each treatment ($\hat{\beta}_{ij} + \hat{g}_{ijk} + \hat{b}_{ijp}$). The table is sorted in decreasing order of between-treatment variation ($\hat{\sigma}_g$) and all the estimates include their 90\%-credible intervals.
```{r full_effects_table2, eval=TRUE, include=TRUE, fig.align="center", out.height="\\textheight", out.width="100%", echo=FALSE}
knitr::include_graphics("tables/variation_of_nutrition_effects1.pdf")
```

Supplementary Table S4 (continued)
```{r full_effects_table3, eval=TRUE, include=TRUE, fig.align="center", out.height="\\textheight", out.width="100%", echo=FALSE}
knitr::include_graphics("tables/variation_of_nutrition_effects2.pdf")
```

```{r top_effect_summary,echo=FALSE, cache=FALSE}
library(dplyr)

effect_summary <- readRDS("patient_summary/two_level_effect_summary.rds")

# Most varying effect between patients regardless the treatment

effect_summary$sigma_b <- as.numeric(effect_summary$sigma_b)
effect_summary$personal_effect <- as.numeric(effect_summary$personal_effect)

top_effect_summary <- effect_summary %>%
  group_by(effect) %>% 
  mutate(personal_max = personal_effect[which.max(personal_effect)]) %>% 
  mutate(personal_min = personal_effect[which.min(personal_effect)]) %>% 
  mutate(personal_variation = personal_max - personal_min) %>% 
  ungroup() %>%
  arrange(desc(personal_variation)) %>%
  select(effect, personal_min, personal_max,personal_variation) %>%
  distinct()

# - pretty name lookup
all_effects <- expand.grid(predictor = assumedpredictors_fat_epros$Name, target = assumedtargets$Name)
all_descs <- expand.grid(predictor = assumedpredictors_fat_epros$Description, target = assumedtargets$Description)
effect_desc.df <- data.frame(effect <- paste0(all_effects$predictor,"_",all_effects$target), row.names = NULL)
effect_desc.df$desc <- paste0(effect <- paste0(all_descs$predictor," -> ",all_descs$target))
effect_desc.df$nutrient <- all_descs$predictor
effect_desc.df$concentration <- all_descs$target
colnames(effect_desc.df) <- c("effect", "effect_description", "nutrient", "concentration")

top_effect_summary <- top_effect_summary %>% inner_join(effect_desc.df, by = "effect")

```

\newpage
Supplementary Table S5: Table shows 20 effects that have highest variation between patients by comparing minimum and maximum of expected effect strength.
```{r top_effect_summary_table,echo=FALSE, cache=FALSE}
kable(top_effect_summary[1:20, c(5,2,3)], "latex", escape = T, booktabs = T, row.names = FALSE, digits = 3, col.names = c("Effect", "Min", "Max")) %>%
    kable_styling(latex_options = "HOLD_position") %>%
    kable_styling(latex_options = c("basic", "condensed"), full_width = FALSE) %>%
    add_header_above(c(" " = 1, "Expected personal effect strength" = 2), escape = F) %>%
    row_spec(0,bold=TRUE)
```

```{r recommendation_overview_simulation, eval=FALSE,echo=FALSE, message=FALSE}
source("mebn/v3/MEBNv3.r")
library(igraph)
library(gridExtra)
library(dplyr)
  
# Query parameters
simulation <- "X50_B50_A50_I50"

# - data parameters
graph_dir <- "graphs/mv3_cross_two_levels/"

# - evidence / query parameters
#pk_recommendations <- c(3.4,4.7)
#fppi_recommendations <- c(0.4,1.8)
#palb_recommendations <- c(34,48)
#recommeded_concentrations <- t(array(c(pk_recommendations, fppi_recommendations, palb_recommendations), dim = c(2,3)))

potassium_org_mean <- mean(dialysis$kalium) # 2755
potassium_org_sd <- sd(dialysis$kalium) # 949
  
phosphorous_org_mean <- mean(dialysis$fosfori) # 2755
phosphorous_org_sd <- sd(dialysis$fosfori) # 949

datadesc <- datadesc_fat_epros
uniform_proposaldist_limits <- c(((datadesc[datadesc$Name=="kalium",]$Lowerbound - potassium_org_mean) / potassium_org_sd),
                                ((datadesc[datadesc$Name=="kalium",]$Upperbound - potassium_org_mean) / potassium_org_sd),
                                ((datadesc[datadesc$Name=="fosfori",]$Lowerbound - phosphorous_org_mean) / phosphorous_org_sd),
                                ((datadesc[datadesc$Name=="fosfori",]$Upperbound - phosphorous_org_mean) / phosphorous_org_sd))

# - collect personal statistics in this data frame
patient_summary <- data.frame(matrix(ncol = 25, nrow = 0), row.names = NULL)

colnames(patient_summary) <- c("person_id", "subject_code", "treatment", "accepted", "max_conc_prob", "max_pk_prob", "max_fppi_prob", "max_palb_prob", "potassium_low", "phosphorous_low", "potassium_high", "phosphorous_high","pk_min","pk_max", "fppi_min", "fppi_max", "palb_min", "palb_max", "estimated_pk","true_pk","estimated_fppi","true_fppi","estimated_palb","true_palb","recommandation_probability")

for (person_id in 1:37) {
  
  print(paste0("Simulating patient ", person_id))

  subject_code <- levels(dialysis$potilas)[person_id]
  treatment <- unique(dialysis[dialysis$potilas == subject_code,]$hoitomuoto)
  
  # Normal ranges of concentration are personalized
  personal_info <- head(dialysis[dialysis$potilas == subject_code,],1)
  recommeded_concentrations <- mebn.get_personal_target_guidelines(personal_info, patient_in_dialysis = TRUE)
  
  print(paste0("personal lower limits: ", as.vector(recommeded_concentrations$lower_limits)))
  print(paste0("personal upper limits: ", as.vector(recommeded_concentrations$upper_limits)))
  
  personal_model_dir <- paste0(graph_dir,person_id)
  personal_graph <- read.graph(paste0(personal_model_dir, "/personal_graph.graphml"), "graphml")
  
  # Sample diet proposals and matching concentration predictions
  
  # - repeating personal behaviour for model checking is moved earlier in this notebook
  personal_repeats <- cbind(0,0,0,0,0,0)
  colnames(personal_repeats) <- c("estimated_pk","true_pk","estimated_fppi","true_fppi","estimated_palb","true_palb")
  
  #source("mebn/v3/MEBNv3.r")

  intake_model <- mebn.Query(reaction_graph = personal_graph,
                           graph_dir = personal_model_dir,
                           queried_nodes = c("kalium","fosfori"),
                           query = recommeded_concentrations,
                           proposal_limits = c(((0 - potassium_org_mean) / potassium_org_sd),
                                             ((5800 - potassium_org_mean) / potassium_org_sd),
                                             ((0 - phosphorous_org_mean) / phosphorous_org_sd),
                                             ((2550 - phosphorous_org_mean) / phosphorous_org_sd)),
                           stan_model_file = "diet/multivariate_intake_recommendation.stan",
                           conc_lower_limits = as.vector(recommeded_concentrations$lower_limits),
                           conc_upper_limits = as.vector(recommeded_concentrations$upper_limits),
                           beta_point_est = "mean",
                           param_point_est = "mean",
                           X_sd_coef = 0, # coef = 0 forces to use the point estimate
                           X_point_est = "mean",
                           posterior_samples = 1000,
                           repeat_only = 0,
                           condition_in_repeat = 0)
  
  # Extract all the samples and correlate concentrations and intake with proposal ids
  
  result <- rstan::extract(intake_model)

  pk <- result$concentration[,,1]
  fppi <- result$concentration[,,2]
  palb <- result$concentration[,,3]

  proposals <- 1:dim(pk)[1]
  proposal_samples <- 1:dim(pk)[2]

  concs.df <- data.frame(proposal=rep(proposals, each = length(proposal_samples)), pk=as.vector(t(pk)),fppi=as.vector(t(fppi)),palb=as.vector(t(palb)))
  
  # Get the original scale levels of potassium and potassium from each diet proposal
  #pot <- result$Q[,1] * potassium_org_sd + potassium_org_mean
  #pho <- result$Q[,2] * phosphorous_org_sd + phosphorous_org_mean

  pot <- result$Q_pred[,1] * potassium_org_sd + potassium_org_mean
  pho <- result$Q_pred[,2] * phosphorous_org_sd + phosphorous_org_mean
  
  intake_proposals.df <- data.frame(proposal=proposals, pot=as.vector(pot),pho=as.vector(pho))
  
  #  Calculate probability of each proposal being within the recommended concentration limits.
  #  The calculation is done based on empirical cumulative distribution function (ecdf) for each concentration distribution. 
  
  pk_recommendations <- c(recommeded_concentrations$lower_limits[1], recommeded_concentrations$upper_limits[1])
  fppi_recommendations <- c(recommeded_concentrations$lower_limits[2], recommeded_concentrations$upper_limits[2])
  palb_recommendations <- c(recommeded_concentrations$lower_limits[3], recommeded_concentrations$upper_limits[3])

  proposal_probs.df <- concs.df %>%
  select(proposal,pk,fppi,palb) %>%
  group_by(proposal) %>%
  summarise(
    pk_prob = (ecdf(pk)(pk_recommendations[2]) - ecdf(pk)(pk_recommendations[1])),
    fppi_prob = (ecdf(fppi)(fppi_recommendations[2]) - ecdf(fppi)(fppi_recommendations[1])),
    palb_prob = (ecdf(palb)(palb_recommendations[2]) - ecdf(palb)(palb_recommendations[1])),
    lowest_conc_prob = min(pk_prob, fppi_prob, palb_prob),
    iid_prob = pk_prob * fppi_prob * palb_prob,
  )
  
  # Filter all accepted diet proposals with a fixed probability limit. 
  acceptance_probability <- 0.50
  
  #   This is denoted with constant 'c' in Algorithm 1.
  #recommandation_probability <- 0.90

  recommandation_probability_levels <- c(0.90, 0.80)
  
  accepted <- 0

  # conc_limits <- concs.df %>%
  #   inner_join(proposal_probs.df, by="proposal") %>%
  #   select(pk,fppi,palb) %>%
  #   summarise(
  #             pk_low = min(pk), 
  #             fppi_low = min(fppi), 
  #             palb_low = min(palb),
  #             pk_high = max(pk), 
  #             fppi_high = max(fppi), 
  #             palb_high = max(palb)
  #   )

  # Find the lowest and highest reachable concentrations  
  # TODO: filter by probability (Above)?
  
  conc_limits <- concs.df %>%
  select(pk,fppi,palb) %>%
  summarise(
            pk_low = min(pk), 
            fppi_low = min(fppi), 
            palb_low = min(palb),
            pk_high = max(pk), 
            fppi_high = max(fppi), 
            palb_high = max(palb)
  )

  # This requires all concentrations from one proposal to be accepted
  # distribution of accepted proposals    

  accepted_proposals.df <- proposal_probs.df %>%
    filter(lowest_conc_prob >= acceptance_probability) %>%
    as.data.frame()

  print(paste0("Proposals accepted ", nrow(accepted_proposals.df)))
  
  # Join intake and resulting concentrations
  accepted_intake.df <- accepted_proposals.df %>% inner_join(intake_proposals.df, by="proposal")

  # Highest probability that can be reached  
  max_conc_prob <- max(proposal_probs.df$lowest_conc_prob)

  # Search for the widest nutrient intake range within the accepted proposals  

  accepted <- 0
  min_pot <- -1
  min_pho <- -1 
  max_pot <- -1
  max_pho <- -1

  accepted_recommandation_probability <- 0

  # - Do we have any accepted proposals?
  if (nrow(accepted_intake.df) > 0)
  {
    saveRDS(accepted_intake.df, paste0("patient_summary/mv3_cross_two_levels/simulation",simulation,"/details/accepted_intake_",person_id,".rds"))

    # Loop possible recommandation_probability_levels and fallback to lower level if higher returns empty
    
    for (recommandation_probability in recommandation_probability_levels)
    {
      # Filter recommendation distributions
      recommendation_intake.df <- accepted_intake.df %>%
        filter(lowest_conc_prob >= recommandation_probability) %>%
        as.data.frame()
      
      print(paste0("Proposals narrowed to ", nrow(recommendation_intake.df)))
  
     if (nrow(recommendation_intake.df) > 0)
     {
        accepted <- 1
        accepted_recommandation_probability <- recommandation_probability
  
        # Pick 2.5% and 97.5% quantiles of recommendation posteriors as recommendation limits   
        pot_quantiles <- quantile(recommendation_intake.df$pot, probs = c(0.025,0.975))
        pho_quantiles <- quantile(recommendation_intake.df$pho, probs = c(0.025,0.975))
    
        # And then closest sampled proposals matching those limits
        min_pot_row <- recommendation_intake.df[which.min(abs(recommendation_intake.df$pot-pot_quantiles[1])),]
        max_pot_row <- recommendation_intake.df[which.min(abs(recommendation_intake.df$pot-pot_quantiles[2])),]
        min_pho_row <- recommendation_intake.df[which.min(abs(recommendation_intake.df$pho-pho_quantiles[1])),]
        max_pho_row <- recommendation_intake.df[which.min(abs(recommendation_intake.df$pho-pho_quantiles[2])),]
    
        # - and nutrient levels
        min_pot <- as.numeric(min_pot_row$pot)
        max_pot <- as.numeric(max_pot_row$pot)
        min_pho <- as.numeric(min_pho_row$pho)
        max_pho <- as.numeric(max_pho_row$pho)
        
        # - concentrations matching those proposals    
        concs_minmax.df <- concs.df[concs.df$proposal %in% c(min_pot_row$proposal,max_pot_row$proposal,min_pho_row$proposal,max_pho_row$proposal),]
        
        saveRDS(concs_minmax.df, paste0("patient_summary/mv3_cross_two_levels/simulation",simulation,"/details/concs_minmax_",person_id,".rds"))
        
        # Break out from loop with this probability
        break
     } 
      else
     {
        accepted <- 0
        min_pot <- -1
        min_pho <- -1 
        max_pot <- -1
        max_pho <- -1
     }
    }
  }

  # Here the min and max concentrations match the nutrient recommendations
  personal_intake <- cbind(person_id,
                                   subject_code,
                                   treatment,
                                   accepted,
                                   max_conc_prob,
                                   max_pk_prob = max(proposal_probs.df$pk_prob),
                                   max_fppi_prob = max(proposal_probs.df$fppi_prob),
                                   max_palb_prob = max(proposal_probs.df$palb_prob),
                                   min_pot, 
                                   min_pho, 
                                   max_pot, 
                                   max_pho,
                                   accepted_recommandation_probability)

  personal_recommendation <- cbind(personal_intake, conc_limits, personal_repeats)
  patient_summary <- rbind(patient_summary,personal_recommendation)

  saveRDS(personal_recommendation, paste0("patient_summary/mv3_cross_two_levels/simulation",simulation,"/overviews/personal_recommendation_",person_id,".rds"))
} 

saveRDS(patient_summary, paste0("patient_summary/mv3_cross_two_levels/simulation",simulation,"/recommendation_summary_posterior_params.rds"))

```

```{r personal_effect_summary,eval=TRUE,echo=FALSE,warning=FALSE,message=FALSE}
library(igraph)
graph_dir <- "graphs/mv3_cross_two_levels/"

personal_effect_summary <- data.frame(matrix(ncol = 23, nrow = 0), row.names = NULL)

for (person_id in 1:37) {
  
  personal_model_dir <- paste0(graph_dir,person_id)
  personal_graph <- read.graph(paste0(personal_model_dir, "/personal_graph.graphml"), "graphml")

  nodes <- V(personal_graph)
  personal_effects <- c(person_id,
                      nodes["kalium"]$value_mean,
                      nodes["fosfori"]$value_mean,
                      nodes["pk"]$value_mean,
                      nodes["fppi"]$value_mean,
                      nodes["palb"]$value_mean,
                      nodes["personal_kalium_pk"]$value,
                      nodes["personal_kalium_pk"]$value_lCI,
                      nodes["personal_kalium_pk"]$value_uCI,
                      nodes["personal_kalium_fppi"]$value,
                      nodes["personal_kalium_fppi"]$value_lCI,
                      nodes["personal_kalium_fppi"]$value_uCI,
                      nodes["personal_kalium_palb"]$value,
                      nodes["personal_kalium_palb"]$value_lCI,
                      nodes["personal_kalium_palb"]$value_uCI,
                      nodes["personal_fosfori_pk"]$value,
                      nodes["personal_fosfori_pk"]$value_lCI,
                      nodes["personal_fosfori_pk"]$value_uCI,
                      nodes["personal_fosfori_fppi"]$value,
                      nodes["personal_fosfori_fppi"]$value_lCI,
                      nodes["personal_fosfori_fppi"]$value_uCI,
                      nodes["personal_fosfori_palb"]$value,
                      nodes["personal_fosfori_palb"]$value_lCI,
                      nodes["personal_fosfori_palb"]$value_uCI)
  
    personal_effect_summary <- rbind(personal_effect_summary,personal_effects)
}

colnames(personal_effect_summary) <- c("person_id","kalium","fosfori","pk","fppi","palb","personal_kalium_pk","personal_kalium_pk_lCI","personal_kalium_pk_uCI","personal_kalium_fppi","personal_kalium_fppi_lCI","personal_kalium_fppi_uCI","personal_kalium_palb","personal_kalium_palb_lCI","personal_kalium_palb_uCI","personal_fosfori_pk","personal_fosfori_pk_lCI","personal_fosfori_pk_uCI","personal_fosfori_fppi","personal_fosfori_fppi_lCI","personal_fosfori_fppi_uCI","personal_fosfori_palb","personal_fosfori_palb_lCI","personal_fosfori_palb_uCI")

#personal_effect_summary
```


```{r concentration_highs_and_lows, echo=FALSE,warning=FALSE,message=FALSE}
library(dplyr)

simulation <- "X50_B50_A50_I50"
#simulation <- "X50_B50_A50_I50_fppi2"
#simulation <- "X5_B5_A50_I50"
#simulation <- "X95_B95_A50_I50"

recom_summary <- readRDS(paste0("patient_summary/mv3_cross_two_levels/simulation",simulation,"/recommendation_summary_posterior_params.rds"))

# Calculate concentration level that matches the recommendation

patient_summary <- data.frame(matrix(ncol = 27, nrow = 0), row.names = NULL)

# colnames(patient_summary) <- c("person_id", "subject_code", "treatment", "accepted", "max_conc_prob", "potassium_low", "phosphorous_low", "potassium_high", "phosphorous_high","pk_min","pk_max", "fppi_min", "fppi_max", "palb_min", "palb_max", "estimated_pk","true_pk","estimated_fppi","true_fppi","estimated_palb","true_palb","pk_low_median","pk_high_median","fppi_low_median","fppi_high_median","palb_low_median","palb_high_median")

for (person_id in 1:37) {
  
  #conf <- 0.90
  conf <- as.numeric(recom_summary[recom_summary$person_id == person_id,]$accepted_recommandation_probability)
  low_conf <- 1 - conf
  
  personal_recommendation <- readRDS(paste0("patient_summary/mv3_cross_two_levels/simulation",simulation,"/overviews/personal_recommendation_",person_id,".rds"))
  
  concs_file <- paste0("patient_summary/mv3_cross_two_levels/simulation",simulation,"/details/concs_minmax_",person_id,".rds")
  
  if (file.exists(concs_file))
  {
    personal_concs <- readRDS(concs_file)
    quantile_limits <- c(low_conf,conf)	

    conc_quants <- personal_concs %>%
      select(proposal,pk,fppi,palb) %>%
      group_by(proposal) %>%
      summarise(quant = quantile_limits,
            pk = quantile(pk, quantile_limits), 
            fppi = quantile(fppi, quantile_limits), 
            palb = quantile(palb, quantile_limits),
            .groups = "keep")
      
    pk_low_rec <- min(conc_quants[conc_quants$quant == low_conf,]$pk)
    pk_high_rec <- max(conc_quants[conc_quants$quant == conf,]$pk)

    fppi_low_rec <- min(conc_quants[conc_quants$quant == low_conf,]$fppi)
    fppi_high_rec <- max(conc_quants[conc_quants$quant == conf,]$fppi)

    palb_low_rec <- min(conc_quants[conc_quants$quant == low_conf,]$palb)
    palb_high_rec <- max(conc_quants[conc_quants$quant == conf,]$palb)

    conc_recs <- data.frame(pk_low_rec,pk_high_rec,fppi_low_rec,fppi_high_rec,palb_low_rec,palb_high_rec)
    colnames(conc_recs) <- c("pk_low_rec","pk_high_rec","fppi_low_rec","fppi_high_rec","palb_low_rec","palb_high_rec")
  }
  else
  {
    conc_recs <- data.frame(0,0,0,0,0,0)
    colnames(conc_recs) <- c("pk_low_rec","pk_high_rec","fppi_low_rec","fppi_high_rec","palb_low_rec","palb_high_rec")
  }
  
  patient_summary <- rbind(patient_summary,c(personal_recommendation,conc_recs))
}

```

```{r muq0_summary, echo=FALSE}
library(igraph)
graph_dir <- "graphs/mv3_cross_two_levels/"

muq0_summary <- data.frame(matrix(ncol = 3, nrow = 0), row.names = NULL)

for (person_id in 1:37) {
  
  personal_model_dir <- paste0(graph_dir,person_id)
  personal_graph <- read.graph(paste0(personal_model_dir, "/personal_graph.graphml"), "graphml")
  
  X <- mebn.GetEvidence(personal_graph, c("kalium", "fosfori"), point_est = "mean")
  Y_int <- mebn.GetIntercept(personal_graph, personal_model_dir, point_est = "mean")
  Beta <- mebn.GetBeta(personal_graph, personal_model_dir, point_est = "mean")
  
  pk_mu_q0 <- Y_int[1] + X$evidence[-X$cond_index,1] %*% Beta[1, -X$cond_index]
  fppi_mu_q0 <- Y_int[2] + X$evidence[-X$cond_index,1] %*% Beta[2, -X$cond_index]
  palb_mu_q0 <- Y_int[3] + X$evidence[-X$cond_index,1] %*% Beta[3, -X$cond_index]
  
  if (pk_mu_q0 < 0) pk_mu_q0 <- 0
  if (fppi_mu_q0 < 0) fppi_mu_q0 <- 0
  if (palb_mu_q0 < 0) palb_mu_q0 <- 0
  
  muq0_summary <- rbind(muq0_summary, c(person_id, pk_mu_q0, fppi_mu_q0, palb_mu_q0))
}

colnames(muq0_summary) <- c("person_id", "pk_mu_q0", "fppi_mu_q0", "palb_mu_q0")
muq0_summary <- round(muq0_summary, 2)
muq0_summary$person_id <- as.character(muq0_summary$person_id)
```

```{r personal_concentration_targets, echo=FALSE}

personal_targets <- data.frame(matrix(ncol = 6, nrow = 0), row.names = NULL)

for (person_id in 1:37) {
  
  subject_code <- levels(dialysis$potilas)[person_id]

  personal_info <- head(dialysis[dialysis$potilas == subject_code,],1)
  recommeded_concentrations <- mebn.get_personal_target_guidelines(personal_info, patient_in_dialysis = TRUE)
  
  lower_limits <- as.vector(recommeded_concentrations$lower_limits)
  upper_limits <- as.vector(recommeded_concentrations$upper_limits)
  
  personal_targets <- rbind(personal_targets, c(person_id, lower_limits, upper_limits))
}

colnames(personal_targets) <- c("person_id", "pk_lower_limit", "fppi_lower_limit", "palb_lower_limit", "pk_upper_limit", "fppi_upper_limit", "palb_upper_limit")
```

# Overview of personal recommendations 
```{r reasoning_plot_ridges, fig.height=8.5, fig.width=8, cache=FALSE,echo=FALSE, message=FALSE, warning=FALSE,fig.cap="The figure shows in two left-most panels personal recommendations of potassium and phosphorous intake ($\\hat{Q}^{min}-\\hat{Q}^{max}$) with predictive distributions of $\\hat{Q}$ illustrated in the background. Each row shows a numeric label of the patient and the type of patient's dialysis treatment (HD = hospital hemodialysis, HHD = home hemodialysis, PD = peritoneal dialysis). Personal recommendations are given for those patients whose resulting plasma concentration levels could be predicted in either over 90\\% probability. Three right-most panels show the estimated concentrations that match the recommended intake. In these panels, grey bars indicate the estimated concentration without the effect of potassium and phosphorous intake, $\\mu_{q0}$. Blue bars indicate the range of concentration that is resulted from modifying phosphorous and potassium intake. It is required that these simulated concentrations stay within the personal target ranges denoted with vertical black lines. The dark blue bar indicates satisfying this requirement in $P^{max} > 90\\%$ confidence and $P^{max} > 80\\%$ confidence would be indicated with light blue, but there aren't any in the results. The best exact probabilities $P^{max}_m$ are shown with percentage figures. Red bars indicate that even the best concentration estimation has $P^{max} < 80\\%$ confidence and personal recommendations are not given for these patients. For these patients, the red and blue bars show the whole reachable ranges. The figure is plotted with ggplot2 package for R language (v 3.3.5, https://ggplot2.tidyverse.org)."}

library(ggridges)
library(ggplot2)
library(grid)
library(gridExtra)

simulation <- "X50_B50_A50_I50"

potassium_org_mean <- mean(dialysis$kalium) # 2755
potassium_org_sd <- sd(dialysis$kalium) # 949
  
phosphorous_org_mean <- mean(dialysis$fosfori) # 2755
phosphorous_org_sd <- sd(dialysis$fosfori) # 949

patient_summary$max_pk_prob <- as.numeric(patient_summary$max_pk_prob)
patient_summary$max_fppi_prob <- as.numeric(patient_summary$max_fppi_prob)
patient_summary$max_palb_prob <- as.numeric(patient_summary$max_palb_prob)

min_accepted_recommandation_probability <- 0.80

accepted_intake.df <- data.frame(matrix(ncol = 9, nrow = 0), row.names = NULL)

for (person_id in 1:37) {
  
  personal_intake_file <- paste0("./patient_summary/mv3_cross_two_levels/simulation",simulation,"/details/accepted_intake_",person_id,".rds")
  
  if (file.exists(personal_intake_file)) {
    
    #personal_recommendation <- paste0("./patient_summary/simulation",simulation,"/overviews/personal_recommendation_",person_id,".rds")

    personal_intake.df <- readRDS(personal_intake_file)
    
    recommendations.df <- personal_intake.df[personal_intake.df$lowest_conc_prob >= min_accepted_recommandation_probability,]
    
    if (nrow(recommendations.df) > 0)
    {
      accepted_intake.df <- rbind(accepted_intake.df, cbind(person_id, recommendations.df))
    } else {
      emptyrow <- as.data.frame(matrix(c(person_id,0,0,0,0,0,0,0,0), ncol = 9, nrow = 1))
      colnames(emptyrow) <- c("person_id","proposal", "pk_prob", "fppi_prob", "palb_prob", "lowest_conc_prob", "iid_prob", "pot", "pho")  
      accepted_intake.df <- rbind(accepted_intake.df, emptyrow)
    }

  } else {
    emptyrow <- as.data.frame(matrix(c(person_id,0,0,0,0,0,0,0,0), ncol = 9, nrow = 1))
    colnames(emptyrow) <- c("person_id","proposal", "pk_prob", "fppi_prob", "palb_prob", "lowest_conc_prob", "iid_prob", "pot", "pho")  
    accepted_intake.df <- rbind(accepted_intake.df, emptyrow)
  }
}

# intake reasoning

patient_summary$person_id <- as.character(patient_summary$person_id)
accepted_intake.df$person_id <- as.character(accepted_intake.df$person_id)

# english abbreviations for dialysis treatments
patient_summary[patient_summary$treatment == "OHD",]$treatment <- "HD"
patient_summary[patient_summary$treatment == "KHD",]$treatment <- "HHD"
patient_summary[patient_summary$treatment == "PD",]$treatment <- "PD"

intake_reasoning.df <- accepted_intake.df %>%
    inner_join(patient_summary, by="person_id") %>%
    mutate(max_conc_prob = as.numeric(max_conc_prob)) %>%
    mutate(max_recom_norm = ifelse(accepted == 1, (as.numeric(max_pot) - potassium_org_mean) / potassium_org_sd + (as.numeric(max_pho) - phosphorous_org_mean) / phosphorous_org_sd + 1000, max_conc_prob)) %>%
  mutate(person_text = paste0(person_id, " ",treatment))

#  mutate(min_pot = ifelse(min_pot == "-1", 0, as.numeric(min_pot))) %>%
#  mutate(max_pot = ifelse(max_pot == "-1", 2500, as.numeric(max_pot))) %>%
#  mutate(min_pho = ifelse(min_pho == "-1", 0, as.numeric(min_pho))) %>%
#  mutate(max_pho = ifelse(max_pho == "-1", 1000, as.numeric(max_pho))) %>%
  
intake_reasoning.df$pot_text <- paste0(round(as.numeric(intake_reasoning.df$min_pot),0)," - ", round(as.numeric(intake_reasoning.df$max_pot),0))
intake_reasoning.df$pho_text <- paste0(round(as.numeric(intake_reasoning.df$min_pho),0)," - ", round(as.numeric(intake_reasoning.df$max_pho),0))

intake_reasoning.df[intake_reasoning.df$min_pho == "-1",]$pot_text <- ""
intake_reasoning.df[intake_reasoning.df$min_pho == "-1",]$pho_text <- ""
                    
# concentration reasoning

personal_effect_summary$person_id <- as.character(personal_effect_summary$person_id)
muq0_summary$person_id <- as.character(muq0_summary$person_id)
personal_targets$person_id <- as.character(personal_targets$person_id)

# TODO: Tallenna myös osittain hyväksytyt pitoisuudet jotta tämä toimisi. näyttäisi kivemmalta.
# nyt _recejä ei ole tallessa

#  mutate(pk_low_rec = ifelse(max_pk_prob >= 0.8, pk_low_rec, pk_low)) %>%
#  mutate(pk_high_rec = ifelse(max_pk_prob >= 0.8, pk_high_rec, pk_high)) %>%

conc_reasoning.df <- patient_summary %>%
  inner_join(personal_effect_summary, by="person_id") %>%
  inner_join(muq0_summary, by="person_id") %>%
  inner_join(personal_targets, by="person_id") %>%
  mutate(accepted = as.numeric(accepted)) %>%
  mutate(max_conc_prob = as.numeric(max_conc_prob)) %>%
  mutate(pk_low_rec = ifelse(accepted == 1, pk_low_rec, pk_low)) %>%
  mutate(pk_high_rec = ifelse(accepted == 1, pk_high_rec, pk_high)) %>%
  mutate(fppi_low_rec = ifelse(accepted == 1, fppi_low_rec, fppi_low_rec)) %>%
  mutate(fppi_high_rec = ifelse(accepted == 1, fppi_high_rec, fppi_high)) %>%
  mutate(palb_low_rec = ifelse(accepted == 1, palb_low_rec, palb_low)) %>%
  mutate(palb_high_rec = ifelse(accepted == 1, palb_high_rec, palb_high)) %>%
  mutate(max_recom_norm = ifelse(accepted == 1, (as.numeric(max_pot) - potassium_org_mean) / potassium_org_sd + (as.numeric(max_pho) - phosphorous_org_mean) / phosphorous_org_sd + 1000, max_conc_prob)) %>%
  mutate(pk_accepted = ifelse(max_pk_prob >= 0.9, "high", ifelse(max_pk_prob >= 0.8, "low", "failed"))) %>%
  mutate(fppi_accepted = ifelse(max_fppi_prob >= 0.9, "high", ifelse(max_fppi_prob >= 0.8, "low", "failed"))) %>%
  mutate(palb_accepted = ifelse(max_palb_prob >= 0.9, "high", ifelse(max_palb_prob >= 0.8, "low", "failed")))

mu_q0_color = "grey"
recom_color = "black"
default_recom_color = "lightgrey"

conc_reasoning.df$pk_prob_text <- paste0(round(conc_reasoning.df$max_pk_prob * 100, 0), "%")
conc_reasoning.df$fppi_prob_text <- paste0(round(conc_reasoning.df$max_fppi_prob * 100, 0), "%")
conc_reasoning.df$palb_prob_text <- paste0(round(conc_reasoning.df$max_palb_prob * 100, 0), "%")
conc_reasoning.df$treatment <- factor(conc_reasoning.df$treatment)

# Special case: No valid combination found for accepted concentrations

no_valid <- conc_reasoning.df %>%
  filter(accepted == 0 & max_pk_prob >= min_accepted_recommandation_probability & max_fppi_prob >= min_accepted_recommandation_probability & max_palb_prob >= min_accepted_recommandation_probability) %>%
  mutate(pk_prob_text = paste0("<",round(as.numeric(max_conc_prob) * 100, 0), "%")) %>%
  mutate(fppi_prob_text = "") %>%
  mutate(palb_prob_text = "") %>%
  mutate(pk_accepted = "failed") %>%
  mutate(fppi_accepted = "failed") %>%
  mutate(palb_accepted = "failed")

conc_reasoning.df[conc_reasoning.df$person_id %in% as.vector(no_valid$person_id),] <- no_valid

ProbColorPalette <- c("#F44336","#4CAF50","#81C784")
ProbColorPalette <- c("#F44336","#0000CD","#00BFFF") # <80%, 90%, 80%

#      geom_text(stat = "unique", aes(x=2500, label=pot_text), size=2.5, width=0.6) +

pot_plot <- ggplot(intake_reasoning.df, aes(x = pot, y = reorder(person_text, max_recom_norm + as.numeric(accepted_recommandation_probability) * 100000), group = person_id)) +
      geom_density_ridges(aes(rel_min_height=0.01, scale = 1.0), fill = "#00BFC4", color = "#009FA4") +
      theme_bw() +
      scale_x_continuous(breaks=c(0,2500,5800), limits = c(0,5850)) +
      geom_text(stat = "unique", aes(x=2500, label=pot_text), size=2.5) +
      geom_vline(xintercept = 2500, linetype="solid", color = "black", size=0.2) +
      labs(x = element_blank(), y = "Patients") +
      theme(axis.text.x = element_text(size=6),axis.title = element_text(size=9), axis.text = element_text(size=7), plot.title = element_text(size=9)) +
      ggtitle("Pot. (mg/d)")

pho_plot <- ggplot(intake_reasoning.df, aes(x = pho, y = reorder(person_id, max_recom_norm + as.numeric(accepted_recommandation_probability) * 100000), group = person_id)) +
      geom_density_ridges(aes(rel_min_height=0.01, scale = 1.3), fill = "#00BFC4", color = "#009FA4") +
      theme_bw() +
      scale_x_continuous(breaks=c(0,1000,2550), limits = c(0,3000)) +
      geom_text(stat = "unique", aes(x=1000, label=pho_text), size=2.5) +
      geom_vline(xintercept = 1000, linetype="solid", color = "black", size=0.2) +
      labs(x = element_blank(), y = element_blank()) +
      theme(axis.text.x = element_text(size=6),axis.title = element_text(size=9), axis.text.y = element_blank(), plot.title = element_text(size=9)) +
      ggtitle("Phosp. (mg/d)")

pk_plot <- ggplot(data = conc_reasoning.df, aes(y = reorder(person_id, max_recom_norm + as.numeric(accepted_recommandation_probability) * 100000)), group = accepted) +
      geom_bar(stat='identity', aes(x=pk_mu_q0), fill=mu_q0_color, color=mu_q0_color, width=0.6, show.legend = FALSE) +
      geom_bar(stat='identity', aes(x=pk_high_rec, fill=pk_accepted), width=0.3, show.legend = FALSE) +
      geom_bar(stat='identity', aes(x=pk_low_rec), fill=mu_q0_color, color=mu_q0_color, width=0.6, show.legend = FALSE) +
      geom_text(aes(x=1, label=pk_prob_text), size=2.5, fontface="bold", color="white") +
      theme_bw() +
      scale_color_manual(aesthetics = "fill", values = ProbColorPalette) + 
      scale_x_continuous(breaks=c(0,3.4,4.7,6)) +
      geom_text(aes(x=pk_lower_limit, label="|"), fontface="plain", size=4, color="black") +
      geom_text(aes(x=pk_upper_limit, label="|"), fontface="plain", size=4, color="black") +
      #geom_point(aes(x=pk_lower_limit), shape=3) +
      #geom_point(aes(x=pk_upper_limit), shape=3) +
      labs(x = element_blank(), y = element_blank()) +
      theme(axis.text.y = element_blank(), axis.text.x = element_text(size=6), axis.title = element_text(size=9), plot.title = element_text(size=9)) +
      ggtitle("P-K (mmol/l)")

fppi_plot <- ggplot(data = conc_reasoning.df, aes(y = reorder(person_id, max_recom_norm + as.numeric(accepted_recommandation_probability) * 100000)), group = accepted) +
      geom_bar(stat='identity', aes(x=fppi_mu_q0), fill=mu_q0_color, color=mu_q0_color, width=0.6, show.legend = FALSE) +
      geom_bar(stat='identity', aes(x=fppi_high_rec, fill=fppi_accepted), width=0.3, show.legend = FALSE) +
      geom_bar(stat='identity', aes(x=fppi_low_rec), fill=mu_q0_color, color=mu_q0_color, width=0.6, show.legend = FALSE) +
      geom_text(aes(x=0.5, label=fppi_prob_text), fontface="bold", size=2.5, color="white") +
      scale_x_continuous(breaks=c(0,0.9,1.78,3)) +
      geom_text(aes(x=fppi_lower_limit, label="|"), fontface="plain", size=4, color="black") +
      geom_text(aes(x=fppi_upper_limit, label="|"), fontface="plain", size=4, color="black") +
      #geom_point(aes(x=fppi_lower_limit), shape=3) +
      #geom_point(aes(x=fppi_upper_limit), shape=3) +
      theme_bw() +
      scale_color_manual(aesthetics = "fill", values = ProbColorPalette) + 
      theme(axis.text.y = element_blank(), axis.text.x = element_text(size=6), axis.title = element_text(size=9), plot.title = element_text(size=9)) +
      labs(x = element_blank(), y = element_blank()) +
      ggtitle("fP-Pi (mmol/l)")

palb_plot <- ggplot(data = conc_reasoning.df, aes(y = reorder(person_id, max_recom_norm + as.numeric(accepted_recommandation_probability) * 100000)), group = accepted) +
      geom_bar(stat='identity', aes(x=palb_mu_q0), fill=mu_q0_color, color=mu_q0_color, width=0.6, show.legend = FALSE) +
      geom_bar(stat='identity', aes(x=palb_high_rec, fill=palb_accepted), width=0.3, show.legend = FALSE) +
      geom_bar(stat='identity', aes(x=palb_low_rec), width=0.6, fill=mu_q0_color, color=mu_q0_color, show.legend = FALSE) +
      #geom_text(aes(x=0.5, label=palb_accepted), fontface="bold", size=2.5, color="white") +
      geom_text(aes(x=7, label=palb_prob_text), fontface="bold", size=2.5, color="white") +
      scale_x_continuous(breaks=c(0,36,45)) +
      geom_text(aes(x=palb_lower_limit), label="|", fontface="plain", size=4, color="black") +
      geom_text(aes(x=palb_upper_limit), label="|", fontface="plain", size=4, color="black") +
      #geom_point(aes(x=palb_lower_limit), shape=3) +
      #geom_point(aes(x=palb_upper_limit), shape=3) +
      theme_bw() +
      scale_color_manual(aesthetics = "fill", values = ProbColorPalette) + 
      theme(axis.text.y = element_blank(), axis.text.x = element_text(size=6), axis.title = element_text(size=9), plot.title = element_text(size=9)) +
      labs(x = element_blank(), y = element_blank()) +
      ggtitle("P-Alb (g/l)")

recommendation_plot <- grid.arrange(pot_plot, pho_plot, pk_plot, fppi_plot, palb_plot, ncol=5,nrow=1, padding=0,widths=c(1.32,1,1,1,1))
ggsave(paste0("figures/recommendation_plot_X50_B50.pdf"), plot = recommendation_plot, width = 6, height = 6, scale = 1)

```

## Personal recommendation for a patient

```{r nuts_intake_method_figure, cache=FALSE, eval=TRUE, echo=FALSE,message=FALSE, warning=FALSE, fig.cap="Figure shows detailed intake recommendation for patient 10 from the previous recommendation table. The intake plot on the left shows the posterior samples of diet configurations that result concentrations of plasma potassium (P-K), fasting plasma phosphate (fP-Pi) and plasma albumin (P-Alb) to stay within their recommended limits. These levels are marked with vertical solid lines in the concentration panels. The black point in the middle of intake plot represents the patient's current potassium and phosphorous intake. Current concentrations matching this intake are shown with dashed vertical lines. Reported recommendation is shown with a rectangle that contains $95\\%$ of diet proposals that result recommended concentrations over $90\\%$ accuracy. Colouring of the rectangle sides match the concentration estimates in the right hand panels. The figure is plotted with ggplot2 package for R language (v 3.3.5, https://ggplot2.tidyverse.org)."}
library(ggplot2)
library(gridExtra)
library(ggrepel)

# Plot the personal recommendations
for (person_id in 36:36) {
  
  subject_code <- levels(dialysis$potilas)[person_id]

  personal_info <- head(dialysis[dialysis$potilas == subject_code,],1)
  recommeded_concentrations <- mebn.get_personal_target_guidelines(personal_info, patient_in_dialysis = TRUE)

  pk_recommendations <- c(recommeded_concentrations$lower_limits[1], recommeded_concentrations$upper_limits[1])
  fppi_recommendations <- c(recommeded_concentrations$lower_limits[2], recommeded_concentrations$upper_limits[2])
  palb_recommendations <- c(recommeded_concentrations$lower_limits[3], recommeded_concentrations$upper_limits[3])
    
#  person_id <- 28
  simulation <- "X50_B50_A50_I50"
  
  accepted_intake <- paste0("./patient_summary/mv3_cross_two_levels/simulation",simulation,"/details/accepted_intake_",person_id,".rds")
  
  if (file.exists(accepted_intake)) {
    personal_recommendation <- paste0("./patient_summary/mv3_cross_two_levels/simulation",simulation,"/overviews/personal_recommendation_",person_id,".rds")
    patient_summary <- readRDS(personal_recommendation)
    
    info <- patient_summary[patient_summary$person_id==person_id,]
    min_k <- min(dialysis[dialysis$potilas == info$subject_code,]$kalium)
    max_k <- max(dialysis[dialysis$potilas == info$subject_code,]$kalium)
    avg_k <- mean(dialysis[dialysis$potilas == info$subject_code,]$kalium)
    
    min_f <- min(dialysis[dialysis$potilas == info$subject_code,]$fosfori)
    max_f <- max(dialysis[dialysis$potilas == info$subject_code,]$fosfori)
    avg_f <- mean(dialysis[dialysis$potilas == info$subject_code,]$fosfori)
    
    # these are not anymore retrieved at simulation
    
    true_concs <- dialysis %>%
     filter(potilas == subject_code) %>%
     select(havainto, pk, fppi, palb)

    info$true_pk <- mean(true_concs$pk)
    info$true_fppi <- mean(true_concs$fppi)
    info$true_palb <- mean(true_concs$palb)
    
    accepted_intake.df <- readRDS(accepted_intake)
  
    if (patient_summary$min_pot != -1)
    {
      concs_minmax.df <- readRDS(paste0("patient_summary/mv3_cross_two_levels/simulation",simulation,"/details/concs_minmax_",person_id,".rds"))
    }
    
    rec_file <- paste0("patient_summary/mv3_cross_two_levels/simulation",simulation,"/overviews/personal_recommendation_",person_id,".rds")
    personal_recommendation.df <- readRDS(rec_file)

    MyPalette <- c("#3399FF", "#FF7F0E", "#00BFC4", "#D62728", "#AAAAAA")
    MyPalette <- c("#2860F6", "#FF7F0E", "#00BFC4", "#D62728", "#AAAAAA")
    #MyPalette <- c("#FF0000", "#00FF00", "#0000FF", "#FAC218", "#AAAAAA")
    
    accepted_intake.df$limiting_proposal <- sapply(accepted_intake.df$proposal, function(x) ifelse(x %in% concs_minmax.df$proposal,x,1000000))
    
    if (patient_summary$min_pot == -1)
    {
        intake_plot <- ggplot(accepted_intake.df, aes(x = pot, y = pho)) +
          geom_point(aes(colour=lowest_conc_prob), show.legend = TRUE) +
          labs(x = "Potassium (mg/d)", y = "Phosphorous (mg/d)") +
          annotate("pointrange",x = avg_k, y = avg_f, ymin = min_f, ymax = max_f, xmin = min_k, xmax = max_k, size = 0.1, colour = "black") +
          scale_colour_gradient2(low = "#FFFFFF",
                                 high = "#0000FF",
                                 midpoint = 0.75,
                                 breaks = seq(0.5, 1, by = 0.05),
                                 limits = c(0.5, 1),
                                 labels = as.character(seq(50, 100, by = 5))) +
          scale_x_continuous(limits=c(0,5000), breaks=c(0,1000,2000,3000,4000,5000)) +
          scale_y_continuous(limits=c(0,2500), breaks=c(0,500,1000,1500,2000,2500)) + 
          theme(legend.position="bottom", axis.text.x = element_text(size=6), axis.title = element_text(size=6), plot.title = element_text(size=10)) + 
          guides(colour = guide_colorbar(title = "Probability of reaching target concs.", title.theme = element_text(size=7), title.position = "top", barwidth = unit(80,"mm"))) 
          # geom_label_repel(aes(label = proposal, colour = proposal),
          #                  data = subset(accepted_intake.df, limiting_proposal < 1000000),
          #                  box.padding   = 0.35, 
          #                  point.padding = 0.5,
          #                  segment.color = "grey")
        
        reasoning_plot <- intake_plot
    } 
    else
    {
      lim_proposals <- accepted_intake.df %>% filter(limiting_proposal < 1000000)

      min_pot <- lim_proposals[which.min(lim_proposals$pot),]$proposal
      max_pot <- lim_proposals[which.max(lim_proposals$pot),]$proposal
      min_pho <- lim_proposals[which.min(lim_proposals$pho),]$proposal
      max_pho <- lim_proposals[which.max(lim_proposals$pho),]$proposal

      # Order density colors like this
      
      proposal_order <- data.frame(proposal = c(min_pot, max_pot, min_pho, max_pho))
      proposal_order$color_order <- c(1,2,3,4)
      
      concs_minmax.df <- concs_minmax.df %>% inner_join(proposal_order, by = "proposal")
      
      # Plotting

      intake_plot <- ggplot(subset(accepted_intake.df, lowest_conc_prob >= 0.50), aes(x = pot, y = pho)) +
        geom_point(aes(colour=lowest_conc_prob), show.legend = TRUE) +
        geom_rect(fill = "#FFFFFF00", color = "white", size = 0.6,
                    aes(xmin = as.numeric(personal_recommendation.df$min_pot),
                    ymin = as.numeric(personal_recommendation.df$min_pho),
                    xmax = as.numeric(personal_recommendation.df$max_pot),
                    ymax = as.numeric(personal_recommendation.df$max_pho))) +
        geom_rect(fill = MyPalette[1], color = MyPalette[1], size = 0.6,
                    aes(xmin = as.numeric(personal_recommendation.df$min_pot),
                    ymin = as.numeric(personal_recommendation.df$min_pho),
                    xmax = as.numeric(personal_recommendation.df$max_pot),
                    ymax = as.numeric(personal_recommendation.df$min_pho))) +
        geom_rect(fill = MyPalette[2], color = MyPalette[2], size = 0.6,
                    aes(xmin = as.numeric(personal_recommendation.df$min_pot),
                    ymin = as.numeric(personal_recommendation.df$max_pho),
                    xmax = as.numeric(personal_recommendation.df$max_pot),
                    ymax = as.numeric(personal_recommendation.df$max_pho))) +
        geom_rect(fill = MyPalette[3], color = MyPalette[3], size = 0.6,
                    aes(xmin = as.numeric(personal_recommendation.df$min_pot),
                    ymin = as.numeric(personal_recommendation.df$min_pho),
                    xmax = as.numeric(personal_recommendation.df$min_pot),
                    ymax = as.numeric(personal_recommendation.df$max_pho))) +
        geom_rect(fill = MyPalette[4], color = MyPalette[4], size = 0.6,
                    aes(xmin = as.numeric(personal_recommendation.df$max_pot),
                    ymin = as.numeric(personal_recommendation.df$min_pho),
                    xmax = as.numeric(personal_recommendation.df$max_pot),
                    ymax = as.numeric(personal_recommendation.df$max_pho))) +        
        labs(x = "Potassium (mg/d)", y = "Phosphorous (mg/d)") +
        annotate("pointrange",x = avg_k, y = avg_f, ymin = min_f, ymax = max_f, xmin = min_k, xmax = max_k, size = 0.1, colour = "black") +
        scale_colour_gradient2(low = "#FFFFFF",
                               high = "#0096FF",
                               midpoint = 0.75,
                               breaks = seq(0.5, 1, by = 0.05),
                               limits = c(0.5, 1),
                               labels = as.character(seq(50, 100, by = 5))) +
        scale_x_continuous(limits=c(0,5800), breaks=c(0,1000,2000,3000,4000,5000,5800)) +
        scale_y_continuous(limits=c(0,2550), breaks=c(0,500,1000,1500,2000,2550)) + 
        theme(legend.position="bottom", axis.text = element_text(size=6), axis.title = element_text(size=8), plot.title = element_text(size=10)) + 
        guides(colour = guide_colorbar(title = "Probability % of reaching target concentrations", title.theme = element_text(size=8), label.theme = element_text(size=6), title.position = "top", barwidth = unit(80,"mm"))) 
        # geom_label_repel(aes(label = proposal, colour = proposal),
        #                  data = subset(accepted_intake.df, limiting_proposal < 1000000),
        #                  box.padding   = 0.35, 
        #                  point.padding = 0.5,
        #                  segment.color = "grey")
      
      # what kind of concentration these intakes result?
      pk_plot <- ggplot(concs_minmax.df, aes(x = pk, group=factor(color_order))) +
        geom_density(aes(colour=factor(color_order)), show.legend = FALSE) +
        labs(x = "P-K (mmol/l)", y = NULL, title=element_blank()) +
        geom_vline(xintercept = pk_recommendations, linetype="solid", color = "black", size=0.2) +
        geom_vline(xintercept = as.numeric(info$true_pk), linetype="dashed", color = "black", size=0.5) +
        scale_x_continuous(breaks=c(pk_recommendations,3,4,5,6), limits=c(pk_recommendations,2,3,4,5,6,7)) +
        scale_colour_manual(values = MyPalette) +
        annotate("rect", xmin=-Inf, xmax=pk_recommendations[1], ymax=Inf, ymin=-Inf, alpha = .2) + 
        annotate("rect", xmin=pk_recommendations[2], xmax=Inf, ymax=Inf, ymin=-Inf, alpha = .2) + 
        theme_bw() +
        theme(axis.text.x = element_text(size=6), axis.title = element_text(size=8), plot.title = element_text(size=9)) 
      
      fppi_plot <- ggplot(concs_minmax.df, aes(x = fppi, group=factor(color_order))) +
        geom_density(aes(color=factor(color_order)), show.legend = FALSE) +
        labs(x = "fP-Pi (mmol/l)", y = NULL) +
        geom_vline(xintercept = fppi_recommendations, linetype="solid", color = "black", size=0.2) +
        geom_vline(xintercept = as.numeric(info$true_fppi), linetype="dashed", color = "black", size=0.5) +
        scale_x_continuous(breaks=c(fppi_recommendations,c(0,2,2.5)), limits = c(0.7,2.7)) +
        scale_colour_manual(values = MyPalette) +
        annotate("rect", xmin=-Inf, xmax=fppi_recommendations[1], ymax=Inf, ymin=-Inf, alpha = .2) + 
        annotate("rect", xmin=fppi_recommendations[2], xmax=Inf, ymax=Inf, ymin=-Inf, alpha = .2) + 
        theme_bw() +
        theme(axis.text.x = element_text(size=6), axis.title = element_text(size=8), plot.title = element_text(size=9))
      
      palb_plot <- ggplot(concs_minmax.df, aes(x = palb, color=factor(color_order), group=factor(color_order))) +
        geom_density(aes(color=factor(color_order)), show.legend = FALSE) +
        guides(guide = guide_legend(label.position = "bottom")) +
        labs(x = "P-Alb (g/l)", y = NULL) +
        geom_vline(xintercept = palb_recommendations, linetype="solid", color = "black", size=0.2) +
        geom_vline(xintercept = as.numeric(info$true_palb), linetype="dashed", color = "black", size=0.5) +
        scale_x_continuous(breaks=c(palb_recommendations,30,40), limits = c(30,47)) +
        scale_colour_manual(values = MyPalette) +
        annotate("rect", xmin=-Inf, xmax=palb_recommendations[1], ymax=Inf, ymin=-Inf, alpha = .2) + 
        annotate("rect", xmin=palb_recommendations[2], xmax=Inf, ymax=Inf, ymin=-Inf, alpha = .2) + 
        theme_bw() +
        theme(legend.position="bottom", legend.title = element_blank(), axis.text.x = element_text(size=6), axis.title = element_text(size=8), plot.title = element_text(size=9))
      
      lay <- rbind(c(1,1,2),
                   c(1,1,3),
                   c(1,1,4))
      
      #reasoning_plot <- grid.arrange(intake_plot, pk_plot, fppi_plot, palb_plot, nrow = 2, ncol=2, padding=0) 
      reasoning_plot <- grid.arrange(intake_plot, pk_plot, fppi_plot, palb_plot, layout_matrix = lay, padding=0)
      ggsave(paste0("figures/reasoning_plots/reasoning_plots_",person_id,".pdf"), plot = reasoning_plot, width = 6, height = 6, scale = 1)
    }
    
  } # file exists
}
```

```{r plot_effect_density,eval=FALSE,echo=FALSE}
library(rstan)
library(ggplot2)

#      "verenrasvojenlaakitys" "diabeteslaakitys"      "aktdvit"               "fosforinsitoja"        "renavit"               "hhepros"               "rasvaepros"           
# [8]  "mufaepros"             "pufaepros"             "protepros"             "safaepros"             "kuitu"                 "protgkg"               "sukupuoli"            
# [15] "ekcalkg"               "kalsium"               "fosfori"               "kalium"                "suola"                 "natrium"               "vesi"                 
# [22] "dvit"

# assumedpredictors_fat_epros Name

# This block can be used for inspecting the effect posteriors in detail

#twolvl_fit <- mebn.get_localfit(target_name = "pk_fppi_palb", local_model_cache = "models/BLMM_gamma_mv_cross/two_levels/")

person_id <- 10
subject_code <- "2303"

group_id <- as.numeric(unique(dialysis[dialysis$potilas == subject_code,]$hoitoryhma))

group_id <- 3
                              
pred <- "pufaepros"
idx <- which(assumedpredictors_fat_epros$Name == pred)
palb <- 3

beta <- rstan::extract(twolvl_fit, pars = c(paste0("beta[",palb,",",idx,"]")))
group_effect <- rstan::extract(twolvl_fit, pars = c(paste0("group_effect[",group_id,",",palb,",",idx,"]")))
personal_effect <- rstan::extract(twolvl_fit, pars = c(paste0("personal_effect[",person_id,",",palb,",",idx,"]")))

dens1 <- as.data.frame(beta[[1]])
colnames(dens1) <- c("effect")
dens1$level <- "general"

dens2 <- as.data.frame(group_effect[[1]])
colnames(dens2) <- c("effect")
dens2$level <- "PD_avg"

dens3 <- as.data.frame(personal_effect[[1]])
colnames(dens3) <- c("effect")
dens3$level <- "personal"

dens <- rbind(dens1,dens2,dens3)

m1 <- mean(dens1$effect) 
m2 <- median(dens$effect) 

getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

m3 <- getmode(dens1$effect) 

getmode_continuous <- function(v) {
  d <- density(v)
  d$x[which.max(d$y)]
}

m4 <- getmode_continuous(dens1$effect) # 0.1939

ggplot(dens1, aes(x=effect, fill=level)) +
  geom_density(alpha=.25) +
  geom_vline(xintercept = m1, color = "red") +
  geom_vline(xintercept = m2, color = "green") +
  geom_vline(xintercept = m3, color = "blue") +
  geom_vline(xintercept = m4, color = "yellow") 

```
```{r scatterplot_for_correlations, eval = FALSE, echo = FALSE}

#m4 pufaepros - palb -2.24843

fig_protepros <- ggplot() +
  geom_point(data=dialysis, aes(y = palb, x=protepros, color=potilas), size=0.4, show.legend = FALSE) +
  geom_line(data=dialysis, aes(y = palb, x=protepros, color=potilas), size=0.4, show.legend = FALSE)

fig_protgkg <- ggplot() +
  geom_point(data=dialysis, aes(y = palb, x=protgkg), size=0.4, show.legend = FALSE) +
  geom_line(data=dialysis, aes(y = palb, x=protgkg, color=potilas), size=0.4, show.legend = FALSE) 

fig_pufa <- ggplot() +
  geom_point(data=dialysis, aes(y = palb, x=pufa, color=potilas), size=0.4, show.legend = FALSE) +
  geom_line(data=dialysis, aes(y = palb, x=pufa, color=potilas), size=0.4, show.legend = FALSE)

fig_pufaepros <- ggplot() +
  geom_point(data=dialysis, aes(y = palb, x=mufaepros, color=potilas), size=0.4, show.legend = FALSE) +
  geom_line(data=dialysis, aes(y = palb, x=mufaepros, color=potilas), size=0.4, show.legend = FALSE)

p <- grid.arrange(fig_protepros, fig_protgkg, fig_pufaepros, fig_pufa, nrow = 2, ncol=2, padding=0)

lm(palb ~ protgkg, dialysis) # 1.975
lm(palb ~ protepros, dialysis) # 0.146
lm(palb ~ pufa, dialysis) # 0.02345
lm(palb ~ pufaepros, dialysis) # -0.08943

fig_pufaepros

person_id <- 36
subject_code <- levels(dialysis$potilas)[person_id]
  
dialysis
mean(dialysis[dialysis$potilas == subject_code,]$fosfori)

```

